### 2016-04-15 Rapport/main.tex
rapport template

\documentclass[11pt,a4paper,oneside]{book}
\usepackage{a4wide}                     % Iets meer tekst op een bladzijde
\usepackage[dutch]{babel}               % Voor nederlandstalige hyphenatie (woordsplitsing)
\usepackage{amsmath}                    % Uitgebreide wiskundige mogelijkheden
\usepackage{amssymb}                    % Voor speciale symbolen zoals de verzameling Z, R...
\usepackage{makeidx}                    % Om een index te maken
\usepackage{url}                        % Om url's te verwerken
\usepackage{graphicx}                   % Om figuren te kunnen verwerken
\usepackage[small,bf,hang]{caption}     % Om de captions wat te verbeteren
\usepackage{xspace}                     % Magische spaties na een commando
\usepackage[latin1]{inputenc}           % Om niet ascii karakters rechtstreeks te kunnen typen
\usepackage{float}                      % Om nieuwe float environments aan te maken. Ook optie H!
\usepackage{flafter}                    % Opdat floats niet zouden voorsteken
\usepackage{listings}                   % Voor het weergeven van letterlijke text en codelistings
\usepackage[round]{natbib}              % Voor auteur-jaar citaties.
\usepackage[nottoc]{tocbibind}			% Bibliografie en inhoudsopgave in ToC; zie tocbibind.dvi
\usepackage{eurosym}                    % om het euro symbool te krijgen
\usepackage{textcomp}                   % Voor onder andere graden celsius
\usepackage{fancyhdr}                   % Voor fancy headers en footers
\usepackage[Gray,squaren,thinqspace,thinspace]{SIunits} % Om elegant eenheden te zetten
\usepackage[version=3]{mhchem}          % Voor elegante scheikundige formules

% Volgend package is niet echt nodig. Het laat echter toe om gemakkelijk elektronisch
% te navigeren in je pdf-document. Deze package moet altijd als laatste ingeladen worden.
\usepackage[a4paper,plainpages=false]{hyperref}    % Om hyperlinks te hebben in het pdfdocument.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Algemene instellingen van het document.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% De splitsingsuitzonderingen
\hyphenation{back-slash split-sings-uit-zon-de-ring}

% Het bibliografisch opmaak bestand.
% ZORG ERVOOR DAT bibliodutch.bst ZICH IN JE WERKDIRECTORY BEVINDT!!!
\bibliographystyle{bibliodutch}

\setlength{\parindent}{0cm}             % Inspringen van eerste lijn van paragrafen is niet gewenst.

\renewcommand{\baselinestretch}{1.2} 	% De interlinie afstand wat vergroten.

\graphicspath{{figuren/}}               % De plaars waar latex zijn figuren gaat halen.

\makeindex                              % Om een index te genereren.

\setcounter{MaxMatrixCols}{20}          % Max 20 kolommen in een matrix

% De headers die verschijnen bovenaan de bladzijden, herdefinieren:
\pagestyle{fancy}                       % Om aan te geven welke bladzijde stijl we gebruiken.
\fancyhf{}                              % Resetten van al de fancy settings.
\renewcommand{\headrulewidth}{0pt}      % Geen lijn onder de header. Zet dit op 0.4pt voor een mooie lijn.
\fancyhf[HL]{\nouppercase{\textit{\leftmark}}} % Links in de header zetten we de leftmark,
\fancyhead[HR]{\thepage}                % Rechts in de header het paginanummer.
% Activeer de volgende lijn en desactiveer de vorige om paginanummers onderaan gecentreerd te krijgen.
%\fancyhf[FC]{\thepage}                  % Paginanummers onderaan gecentreerd.

% PDF specifieke opties, niet strict noodzakelijk voor een thesis.
% Is hetgeen verschijnt wanneer je in acroread de documentproperties bekijkt.
\hypersetup{
    pdfauthor = {Mathias Dierickx, Maarten Tindemans, Tim Ranson, Bjorn Vandenbussche},
    pdftitle = {Project computervisie},
    pdfsubject = {},
    pdfkeywords = {}
}

% Het volgende commando zou ervoor moeten zorgen dat er een witte ruimte wordt gelaten tussen
% elke paragraaf. Het zorgt ervoor dat er echter teveel witte ruimte komt boven en onder de
% verschillende titels, gemaakt met \section, subsection...
%%\setlength{\parskip}{0ex plus 0.3ex minus 0.3ex}

% Vandaar dat we expliciet aangeven wanneer we wensen dat een nieuwe paragraaf begint:
% \par zorgt ervoor dat er een nieuwe paragraaf begint en
% \vspace zorgt voor vertikale ruimte.
\newcommand{\npar}{\par \vspace{2.3ex plus 0.3ex minus 0.3ex}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Nieuwe commando's
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% De differentiaal operator
\newcommand{\diff}{\ensuremath{\mathrm{d}}}

% Super en subscript
\newcommand{\supsc}[1]{\ensuremath{^{\text{#1}}}}   % Superscript in tekst
\newcommand{\subsc}[1]{\ensuremath{_{\text{#1}}}}   % Subscript in tekst

% Chemische formule font:
\newcommand{\ch}[1]{\ensuremath{\mathrm{#1}}\xspace}	 
% Chemische pijl naar rechts:
\newcommand{\chpijlr}{\ensuremath{\hspace{1em}\longrightarrow\hspace{1em}}}
% Chemische pijl naar links:
\newcommand{\chpijll}{\ensuremath{\hspace{1em}\longleftarrow\hspace{1em}}}
% Chemische pijl naar links en rechts:
\newcommand{\chpijllr}{\ensuremath{\hspace{1em}\longleftrightarrow\hspace{1em}}}

\newcommand{\vt}[1]{\ensuremath{\boldsymbol{#1}}} % vector in juiste lettertype
\newcommand{\mx}[1]{\ensuremath{\mathsf{#1}}}	  % matrix in juiste lettertype

% Het latex logo in een eenvoudiger commando steken
\newcommand{\latex}{\LaTeX\xspace}

% Het BibTeX logo
\newcommand{\bibtex}{\textsc{Bib}\TeX\xspace}

% Niew commando om bestandnamen anders weer te geven
\newcommand{\bestand}[1]{\lstinline[basicstyle=\sl]{#1}\xspace}

% Niew commando om commando tekst weer te geven
\newcommand{\command}[1]{\lstinline[basicstyle=\tt]{#1}\xspace}
\newcommand{\commandx}[1]{\index{#1}\lstinline[basicstyle=\tt]{#1}\xspace}

%\lstset{morecomment={\%}}
% Commando om latex commando`s weer te geven (x: voor indexing)
%\newcommand{\lcommand}[1]{\lstinline[basicstyle={\tt},{language=[LaTeX]TeX}]{#1}\xspace}
\newcommand{\lcommand}[1]{\lstinline[basicstyle={\tt}]{#1}\xspace}
\newcommand{\lcommandx}[1]{\index{#1}\lstinline[basicstyle=\tt]{#1}\xspace}

% Niew commando om vreemde taal weer te geven (hint: dit commando kan gebruikt
%   worden om latijnse namen, die ook cursief moeten staan, weer te geven.
\newcommand{\engels}[1]{\textit{#1}\xspace}
\newcommand{\engelsx}[1]{\index{#1}\textit{#1}\xspace}

% Niew commando om iets te benadrukken en tegelijkertijd in de index te steken.
\newcommand{\begrip}[1]{\index{#1}\textbf{#1}\xspace}

% Nieuw commando om figuren in te voegen. Gebruik:
% \mijnfiguur[H]{width=5cm}{bestandsnaam}{Het bijschrift bij deze figuur}
\newcommand{\mijnfiguur}[4][ht]{            % Het eerste argument is standaar `ht'.
    \begin{figure}[#1]                      % Beginnen van de figure omgeving
        \begin{center}                      % Beginnen van de center omgeving
            \includegraphics[#2]{#3}        % Het eigenlijk invoegen van de figuur (2: opties, 3: bestandsnaam)
            \caption{#4\label{#3}}          % Het bijschrift (argument 4) en het label (argument 3)
        \end{center}
    \end{figure}
    }

% Nieuw commando om figuren in te voegen. Gebruik:
% \mijnfiguur[H]{bestand-tabular}{Het bijschrift bij deze tabel}    
\newcommand{\mijntabel}[3][ht]{             % Het eerste argument is standaar `ht'.
    \begin{table}[#1]                       % Beginnen van de table omgeving
        \begin{center}                      % Beginnen van de center omgeving
            \caption{#3\label{#2}}          % Het bijschrift (argument 3) en het label (argument 2)
            \input{#2}                      % Invoer van de tabel
        \end{center}
    \end{table}
    }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Nieuwe wiskunde operatoren
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathOperator{\integ}{Integraal}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Nieuwe omgevingen
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Een soort theorem omgeving
\newtheorem{levensles}{Levensles}[chapter]

% Om minder belangrijke delen iets kleiner te zetten.
\newenvironment{MinderBelangrijk}{\small}{}

% Een nieuwe omgeving om letterlijke latex tekst weer te geven.
\lstnewenvironment{llt} 
    {
    \vspace{1.2ex plus 0.5ex minus 0.5ex}   % Beetje ruimte voor de letterlijke tekst
    \lstset{                                % Enkele opties:
        basicstyle={\small\tt},             % Iets kleiner
        %language=[LaTeX]{TeX},              % Syntax highlighting
        stepnumber=0,                       % De lijnen worden niet genummerd
        breaklines=true,                    % Als een lijn te lang is, wordt hij afgebroken
        basewidth={0.5em},                  % Breedte van een letter
        xleftmargin=1em}                    % Inspringing van de linker marge
    }
    {\vspace{0.9ex plus 0.5ex minus 0.5ex}  % Beetje ruimte na de letterlijke tekst
    }

% Een nieuwe omgeving om algemene letterlijke tekst weer te geven.
\lstnewenvironment{lt} 
    {
    \vspace{1.2ex plus 0.5ex minus 0.5ex}   % Beetje ruimte voor de letterlijke tekst
    \lstset{                                % Enkele opties:
        basicstyle={\small\tt},             % Iets kleiner en typmachine lettertype
        stepnumber=0,                       % De lijnen worden niet genummerd
        breaklines=true,                    % Als een lijn te lang is, wordt hij afgebroken
        basewidth={0.5em},                  % Breedte van een letter
        xleftmargin=1em}                    % Inspringing van de linker marge
    }
    {\vspace{0.9ex plus 0.5ex minus 0.5ex}  % Beetje ruimte na de letterlijke tekst
    }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Einde van de preamble.
% Begin van de body:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\frontmatter

\input{titel}                            % Algemene versie

% samenvatting...
%\input{samenvatting}

% De lijnen van de inhoudsopgave iets dichter op elkaar, niet echt nodig voor de thesis, maar 
% voor dit werk kregen we anders een laatste bladzijde met 3 items op.
\renewcommand{\baselinestretch}{1.08} 	% De interlinie afstand wat vergroten.
\small\normalsize                       % Nodig om de baselinestretch goed te krijgen.
\tableofcontents
\renewcommand{\baselinestretch}{1.2} 	% De interlinie afstand wat vergroten.
\small\normalsize                       % Nodig om de baselinestretch goed te krijgen.

\mainmatter

% De verschillende hoofdstukken:
\input{introductie}
\input{besluit}

% De appendices:
%\appendix

%\input{lc-bijlage-mathsymbols}

% De bibliografie en de index
\backmatter

\bibliography{lc-bibliografie}

\end{document}

### 2016-04-16 Dataset/evaluate.py
Update datasets

@@ -1,0 +1,31 @@
+from numpy import *
+import matplotlib.pyplot as plt
+
+seq="01"
+
+gt=genfromtxt("%s/gtdistances.txt"%seq)[:,1]/3.6
+res=genfromtxt("%s/results.txt"%seq)[:,1]/3.6
+speeds=genfromtxt("%s/actualspeeds.txt"%seq)[:,1]
+
+distance=0
+time=0
+for i in range(len(gt)):
+	if res[i]>gt[i]:
+		print("Sorry, you crashed! Speed in frame %d exceeded safety limits."%i)
+		break
+	else:
+		distance+=speeds[i]
+		time+=speeds[i]/res[i]
+	if time>=60:
+		print("You ran out of time after %d frames!"%i)
+		break
+print("You travelled %f metres in %f seconds."%(distance,time))
+
+#plt.figure()
+#plt.plot(gt)
+#plt.plot(speeds)
+#plt.ylim([0,30])
+#plt.show()
+	
+
+

### 2016-04-17 Headers/io.hpp
Added LBP featurevector code

@@ -1,0 +1,67 @@
+#ifndef IO_
+#define IO_
+
+#include <string>
+#include "opencv2/opencv.hpp"
+#include "opencv2/highgui.hpp"
+
+using namespace std;
+using namespace cv;
+
+namespace io {
+
+	/** Output directory **/
+	string dirOutput = "output/";
+
+	/** Shown image properties and helper values **/
+	int imX=50,imY=50
+		,shownImages=0
+		,consecutiveImages=6
+		,normalSize=325;
+
+	void showImage(string windowName,Mat& img, bool resize=true, int imWidth=0, int imHeight=0){
+	    int windowFlag, xSize, ySize, margin=3;
+	    if(resize){
+	        windowFlag = WINDOW_NORMAL;
+	        xSize = ySize = normalSize;
+	    } else {
+	        windowFlag = WINDOW_AUTOSIZE;
+	        xSize = imWidth;
+	        ySize = imHeight;
+	    }
+
+	    namedWindow(windowName,windowFlag);
+	    imshow(windowName, img);
+	    moveWindow(windowName,imX,imY);
+	    imX += xSize+margin;
+	    if(++shownImages%consecutiveImages == 0){
+	    	imY += ySize+margin;
+	    }
+	}
+
+	void saveImage(string fn, Mat& img){
+	    imwrite(dirOutput+fn+".png",img);
+	}
+
+	template <typename T>
+	void printVector(ostream& os, const vector<T>& v){    int i=0;
+	    while(i<v.size()-1){
+	        os<<v[i]<<";";
+	        i++;
+	    }
+	    os<<v[i];
+	}
+
+	void checkDir(string dir){
+	    struct stat st;
+	    const char* cDir = dir.c_str();
+	    if(stat(cDir,&st) == -1){
+	        if(mkdir(cDir, S_IRWXU | S_IRWXG | S_IROTH | S_IXOTH ) != 0){
+	            cerr<<"Error in checkOutputDir : failed to create dir "<<dir<<endl;
+	            throw;
+	        }
+	    }
+	}
+}
+
+#endif

### 2016-04-17 Headers/lbpfeaturevector.hpp

@@ -1,0 +1,196 @@
+#ifndef FEATURE_VECTORS_
+#define FEATURE_VECTORS_
+
+#include <cstdio>
+#include <string>
+#include <fstream>
+#include <vector>
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <unistd.h>
+#include "opencv2/opencv.hpp"
+#include "opencv2/highgui.hpp"
+
+#include "io.hpp"
+
+using namespace std;
+using namespace cv;
+
+class LbpFeatureVector{
+private:
+	/** Block configuration **/
+	int outerMargin;
+	int innerMargin;
+	int blkX;
+	int blkY;
+	int blkSize;
+
+	/** Featurevector configuration **/
+	int lbpRadius;
+	int histBins;
+public:
+	LbpFeatureVector(): outerMargin(1),innerMargin(0),blkSize(32),lbpRadius(1),histBins(16){
+		blkX = outerMargin;
+		blkY = outerMargin;
+		if(outerMargin<lbpRadius) outerMargin = lbpRadius;
+	}
+
+	LbpFeatureVector(int _outerMargin,int _innerMargin,int _blkSize,int _lbpRadius,int _histBins):outerMargin(_outerMargin),innerMargin(_innerMargin),blkSize(_blkSize),lbpRadius(_lbpRadius),histBins(_histBins){
+		blkX = outerMargin;
+		blkY = outerMargin;
+		if(outerMargin<lbpRadius) outerMargin = lbpRadius;
+	}
+
+	/* Calculate LBP values of pixels in src to dst, with radius r.
+	* Only calculate withing given window, measured by source coordinates (x,y) and winSize.
+	* Based on lbp code from https://github.com/bytefish/opencv/tree/master/lbp
+	*/
+	template <typename _Tp>
+	void _LBP(const Mat& src, Mat& dst, int r, int x, int y, int winSize) {
+	    if( x<r || y<r || x+winSize+r>src.cols || y+winSize+r>src.rows){
+	        cerr << "Error in LBP_ : source coordinates and window size make radius exceed image size" << endl;
+	        throw;
+	    }
+	    dst = Mat::zeros(winSize, winSize, CV_8UC1);
+	    for(int i=y;i<y+winSize;i++) {
+	        for(int j=x;j<x+winSize;j++) {
+	            _Tp center = src.at<_Tp>(i,j);
+	            unsigned char code = 0;
[…]

### 2016-04-17 main.cpp

@@ -4,6 +4,8 @@
 #include "opencv2/highgui.hpp"
 #include <vector>
 #include "LineDetection/LineDetection.cpp"
+#include "Headers/lbpfeaturevector.hpp"
+#include "Headers/io.hpp"
 
 using namespace std;
 using namespace cv;
@@ -53,9 +55,6 @@ int main(int argc, char** argv){
 		bool debugLinedetection = true; // wait after each frame and show all intermediate results
 		vector<Vec4i> lines = ld.getLinesFromImage(frames[i], houghVote, drawLines, debugLinedetection);	
 	}
-
-
-
 
 	showBlendedImages(frames, masks);
 

### 2016-04-17 Headers/io.hpp
Added executable to generate trainingsdata

@@ -62,6 +62,33 @@ namespace io {
 	        }
 	    }
 	}
+
+	void read_images(String folder, String regex, vector<Mat>& images) {
+		VideoCapture cap(folder+"/"+regex);
+		images.clear();
+		while( cap.isOpened() )
+		{
+		    Mat img;
+		    if(!cap.read(img)) {
+		    	return;
+		    }
+		   	images.push_back(img);
+		}
+	}
+
+	void showBlendedImages(vector<Mat> & frames, vector<Mat> & masks) {
+		double alpha = 1.0;
+		double beta = 1.0;
+		namedWindow( "Display window", WINDOW_AUTOSIZE );
+		for(int i = 0; i<frames.size(); i++) {
+			Mat dst;
+			if(frames[i].channels()>1) cvtColor(frames[i], frames[i], CV_RGB2GRAY);
+			if(masks[i].channels()>1) cvtColor(masks[i], masks[i], CV_RGB2GRAY);
+			addWeighted( masks[i] , alpha, frames[i], beta, 0.0, dst);
+			imshow( "Display window", dst );  
+			waitKey(0);
+		}
+	}
 }
 
 #endif

### 2016-04-17 Headers/lbpfeaturevector.hpp

@@ -29,17 +29,9 @@ private:
 	int lbpRadius;
 	int histBins;
 public:
-	LbpFeatureVector(): outerMargin(1),innerMargin(0),blkSize(32),lbpRadius(1),histBins(16){
-		blkX = outerMargin;
-		blkY = outerMargin;
-		if(outerMargin<lbpRadius) outerMargin = lbpRadius;
-	}
+	LbpFeatureVector(): outerMargin(1),innerMargin(0),blkSize(32),lbpRadius(1),histBins(16){}
 
-	LbpFeatureVector(int _outerMargin,int _innerMargin,int _blkSize,int _lbpRadius,int _histBins):outerMargin(_outerMargin),innerMargin(_innerMargin),blkSize(_blkSize),lbpRadius(_lbpRadius),histBins(_histBins){
-		blkX = outerMargin;
-		blkY = outerMargin;
-		if(outerMargin<lbpRadius) outerMargin = lbpRadius;
-	}
+	LbpFeatureVector(int _outerMargin,int _innerMargin,int _blkSize,int _lbpRadius,int _histBins):outerMargin(_outerMargin),innerMargin(_innerMargin),blkSize(_blkSize),lbpRadius(_lbpRadius),histBins(_histBins){}
 
 	/* Calculate LBP values of pixels in src to dst, with radius r.
 	* Only calculate withing given window, measured by source coordinates (x,y) and winSize.
@@ -112,13 +104,18 @@ public:
 	    }
 	}
 
-	void processFrame(string fnFrame){
+	void processFrame(string fnFrame, Mat& img){
     	ostringstream strBldr; // Stringbuilder
 
-		/***    Process frame   ***/
+    	/* Initial configuration */
+		blkX = outerMargin;
+		blkY = outerMargin;
+		if(outerMargin<lbpRadius) outerMargin = lbpRadius;
+
+		/*    Process frame   */
 		string frameName = fnFrame.substr(0,fnFrame.size()-4);
 
-		Mat img = imread(fnFrame, CV_LOAD_IMAGE_COLOR);
+		// Mat img = imread(fnFrame, CV_LOAD_IMAGE_COLOR);
 		
 		if(! img.data){
 		    cerr <<  "Could not open or find frame " << frameName << std::endl ;
@@ -144,7 +141,7 @@ public:
 		int totalBlocks = numBlksWidth*numBlksHeight - 1; // -1 Accounts for 0-indexing
[…]

### 2016-04-17 main.cpp

@@ -10,41 +10,15 @@
 using namespace std;
 using namespace cv;
 
-vector<Mat> read_images(String folder, String regex) {
-	VideoCapture cap(folder+"/"+regex);
-	vector<Mat> matrices;
-	while( cap.isOpened() )
-	{
-	    Mat img;
-	    if(!cap.read(img)) {
-	    	return matrices;
-	    }
-		if(img.channels()>1) cvtColor(img, img, CV_RGB2GRAY);
-	   	matrices.push_back(img);
-	}
-	return matrices;
-}
-
-void showBlendedImages(vector<Mat> & frames, vector<Mat> & masks) {
-	double alpha = 1.0;
-	double beta = 1.0;
-	namedWindow( "Display window", WINDOW_AUTOSIZE );
-	for(int i = 0; i<frames.size(); i++) {
-		Mat dst;
-		addWeighted( masks[i] , alpha, frames[i], beta, 0.0, dst);
-		imshow( "Display window", dst );  
-		waitKey(0);
-	}
-}
-
 int main(int argc, char** argv){
 	if(argc <= 1) {
 		cout << "add parameters: dataset folder" << endl;
 		return 1;
 	}	
 
-	vector<Mat> frames = read_images(argv[1],"frame%05d.png");
-	vector<Mat> masks = read_images(argv[1],"mask%05d.png");
+	vector<Mat> frames,masks;
+	io::read_images(argv[1],"frame%05d.png",frames);
+	io::read_images(argv[1],"mask%05d.png",masks);
 
 	// detect lines
 	// TODO: Filter lines in wrong orientation and choose better parameters
@@ -56,7 +30,7 @@ int main(int argc, char** argv){
 		vector<Vec4i> lines = ld.getLinesFromImage(frames[i], houghVote, drawLines, debugLinedetection);	
 	}
 
-	showBlendedImages(frames, masks);
+	io::showBlendedImages(frames, masks);
 
 	waitKey(0);
     destroyAllWindows();

### 2016-04-17 trainingsdata.cpp

@@ -1,0 +1,38 @@
+#include <string>
+#include <stdio.h>
+#include "Headers/lbpfeaturevector.hpp"
+#include "Headers/io.hpp"
+
+using namespace std;
+using namespace cv;
+
+int main (int argc, char** argv){
+	string datasetFolder = "Dataset/01/";
+
+	if(argc == 2)
+		datasetFolder = argv[1];
+
+    /* Training data */
+    int outerMargin(16),innerMargin(128),blkSize(32),lbpRadius(1),histBins(16);
+    LbpFeatureVector fv(outerMargin,innerMargin,blkSize,lbpRadius,histBins);
+    /* */
+
+    /* Actual data *
+    LbpFeatureVector fv;
+    /* */
+
+	// Use every 10th frame
+	vector<Mat> frames;
+	string frameRegex = "frame%04d0.png";
+	io::read_images(datasetFolder,frameRegex,frames);
+
+	for(int i=0;i<frames.size();i++){
+		char fnFrame[14];
+		sprintf(fnFrame,frameRegex.c_str(),i);
+		fv.processFrame(string(fnFrame),frames[i]);
+	}
+
+    // waitKey(0);
+    destroyAllWindows();
+    return 0;
+}

### 2016-04-17 Headers/lbpfeaturevector.hpp
Cleanup

@@ -114,8 +114,6 @@ public:
 
 		/*    Process frame   */
 		string frameName = fnFrame.substr(0,fnFrame.size()-4);
-
-		// Mat img = imread(fnFrame, CV_LOAD_IMAGE_COLOR);
 		
 		if(! img.data){
 		    cerr <<  "Could not open or find frame " << frameName << std::endl ;

### 2016-04-17 trainingsdata.cpp

@@ -32,7 +32,6 @@ int main (int argc, char** argv){
 		fv.processFrame(string(fnFrame),frames[i]);
 	}
 
-    // waitKey(0);
     destroyAllWindows();
     return 0;
 }

### 2016-06-11 Headers/LineDetection.hpp
Optimized parameters for edge detection

@@ -17,15 +17,19 @@ class LineDetection {
 
 public:
     Mat getLinesFromImage(Mat &image) {
+        return getLinesFromImage(image, false);
+    }
+
+    Mat getLinesFromImage(Mat &image, bool showSteps) {
 
         // parameters
-        int cannyThreshold1 = 60;
-        int cannyThreshold2 = 140;
-        int threshThreshold = 128;
+        int cannyLowerThreshold = 80;
+        int cannyUpperThreshold = cannyLowerThreshold * 2;
+        int threshThreshold = 170;
         int threshMaxval = 255;
-        int houghPMinLineLength = 1;
+        int houghPMinLineLength = 10;
         int houghPGap = 20;
-        int houghPMinVote = 5;
+        int houghPMinVote = 30;
         double houghPDeltaRho = 1;
         double houghPDeltaTheta = PI/180;
 
@@ -35,7 +39,7 @@ public:
 
         // Canny algorithm
         Mat contours;
-        Canny(imgROI, contours, cannyThreshold1, cannyThreshold2);
+        Canny(imgROI, contours, cannyLowerThreshold, cannyUpperThreshold, 3);
         Mat contoursInv;
         threshold(contours, contoursInv, threshThreshold, threshMaxval, THRESH_BINARY_INV);
 
@@ -46,14 +50,17 @@ public:
         HoughLinesP(contours,lines,houghPDeltaRho,houghPDeltaTheta,houghPMinVote, houghPMinLineLength, houghPGap);
         drawDetectedLines(houghP, lines);
 
-        // Uncomment to test parameters
-        /*
-        namedWindow( "Source", CV_WINDOW_AUTOSIZE );
-        imshow( "Source", image );
-        namedWindow( "Lines", CV_WINDOW_AUTOSIZE );
-        imshow( "Lines", houghP );
-        waitKey(0);
-        */
+
+        if(showSteps) {
+            namedWindow( "Input", CV_WINDOW_AUTOSIZE );
+            imshow( "Input", image );
+            namedWindow( "Canny", CV_WINDOW_AUTOSIZE );
+            imshow( "Canny", contours );
+            namedWindow( "threshold", CV_WINDOW_AUTOSIZE );
+            imshow( "threshold", contoursInv );
+            namedWindow( "HoughLinesP", CV_WINDOW_AUTOSIZE );
+            imshow( "HoughLinesP", houghP );
+        }
 
         return houghP;
     };
[…]

### 2016-06-11 main.cpp

@@ -1,48 +1,0 @@
-#include <cstdio>
-#include <string>
-#include "opencv2/opencv.hpp"
-#include "opencv2/highgui.hpp"
-#include <vector>
-#include "LineDetection/LineDetection.cpp"
-#include "Headers/lbpfeaturevector.hpp"
-#include "Headers/io.hpp"
-
-using namespace std;
-using namespace cv;
-
-int main(int argc, char** argv){
-	if(argc <= 1) {
-		cout << "add parameters: dataset folder" << endl;
-		return 1;
-	}	
-
-	vector<Mat> frames,masks;
-	io::read_images(argv[1],"frame%05d.png",frames);
-	io::read_images(argv[1],"mask%05d.png",masks);
-
-	// detect lines
-	LineDetection ld;
-	int initialHoughVote = 150;
-	int initialHoughVote2 = 150;
-	int houghVote = initialHoughVote;
-	int houghVote2 = initialHoughVote2;
-	vector<Mat> lineContours;
-	for(int i=0; i < frames.size(); i++) {
-		bool drawLines = true; // draw detected lines on source image
-		bool debugLinedetection = false; // wait after each frame and show all intermediate results
-		Mat lineContour = ld.getLinesFromImage(frames[i], initialHoughVote, houghVote,initialHoughVote2, houghVote2, drawLines, debugLinedetection);
-		lineContours.push_back(lineContour);
-		namedWindow("LineContours");
-		imshow("LineContours",lineContours);
-		waitKey(0);
-		destroyAllWindows();
-
-	}
-
-	io::showBlendedImages(frames, masks);
-
-	waitKey(0);
-    destroyAllWindows();
-	return 0;
-}																																																																																							
-

### 2016-06-11 maxSpeed.cpp

@@ -12,7 +12,7 @@
 #include <dirent.h>
 #include <iomanip>
 
-#include "Headers/LineDetection.cpp"
+#include "Headers/LineDetection.hpp"
 #include "Headers/lbpfeaturevector.hpp"
 #include "Headers/io.hpp"
 #include "Headers/svm.hpp"
@@ -237,7 +237,7 @@ void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames
         }
         
         if(laagsteSnelheid!= 90)
-            laagsteSnelheid -=2;
+            laagsteSnelheid -=1;
         
         stringstream ss;
         ss << "Max speed (zelf): " << laagsteSnelheid << " km/u, gtdistances: " << speeds[i] << " km/u";

### 2016-06-11 testLineDetection.cpp

@@ -1,0 +1,39 @@
+#include <cstdio>
+#include <string>
+#include "opencv2/opencv.hpp"
+#include "Headers/io.hpp"
+#include "Headers/LineDetection.hpp"
+#include <iomanip>
+
+
+using namespace std;
+using namespace cv;
+
+int main(int argc, char** argv){
+	if(argc <= 1) {
+		cout << "add parameters: dataset folder" << endl;
+		return 1;
+	}	
+
+	vector<Mat> frames,masks;
+	io::readImages(argv[1],"frame%05d.png",frames);
+	io::readImages(argv[1],"mask%05d.png",masks);
+
+	// detect lines
+	LineDetection ld;
+	for(int i=0; i < frames.size(); i++) {
+		Mat lines = ld.getLinesFromImage(frames[i], false);
+		//waitKey(0);
+		io::checkDir("linetest");
+		stringstream ssOut;
+		ssOut << "linetest/frame" << setfill('0') << std::setw(5) << i << ".png";
+		imwrite(ssOut.str(),lines);
+
+		destroyAllWindows();
+	}
+
+	waitKey(0);
+    destroyAllWindows();
+	return 0;
+}																																																																																							
+

### 2016-06-12 Headers/LineDetection.hpp
Dilate and erode to ignore lines and cobbles on the road

@@ -22,14 +22,16 @@ public:
 
     Mat getLinesFromImage(Mat &image, bool showSteps) {
 
+
+
         // parameters
-        int cannyLowerThreshold = 80;
+        int cannyLowerThreshold = 28;
         int cannyUpperThreshold = cannyLowerThreshold * 2;
-        int threshThreshold = 170;
+        int threshThreshold = 240;
         int threshMaxval = 255;
-        int houghPMinLineLength = 10;
-        int houghPGap = 20;
-        int houghPMinVote = 30;
+        int houghPMinLineLength = 60;
+        int houghPGap = 40;
+        int houghPMinVote = 5;
         double houghPDeltaRho = 1;
         double houghPDeltaTheta = PI/180;
 
@@ -37,32 +39,51 @@ public:
         Rect roi(0, image.cols / 4, image.cols - 1, image.rows - image.cols / 4);
         Mat imgROI = image(roi);
 
+
+        Mat tmp;
+        Size size;
+        size.width  = 90;
+        size.height = 15;
+        Mat kernel = getStructuringElement(MORPH_ELLIPSE, size);
+        //erode verticaal
+        erode(imgROI,tmp,kernel);
+
+        size.width  = 60;
+        size.height = 10;
+        kernel = getStructuringElement(MORPH_ELLIPSE, size);
+        dilate(tmp,tmp,kernel);
+
+
         // Canny algorithm
         Mat contours;
-        Canny(imgROI, contours, cannyLowerThreshold, cannyUpperThreshold, 3);
+        Mat result;
+        Canny(tmp, contours, cannyLowerThreshold, cannyUpperThreshold, 3);
         Mat contoursInv;
-        threshold(contours, contoursInv, threshThreshold, threshMaxval, THRESH_BINARY_INV);
+        //threshold(contours, contoursInv, threshThreshold, threshMaxval, THRESH_BINARY_INV);
+        contours.copyTo(result);
 
         // Do probabilistic Hough
         Mat houghP(imgROI.size(), CV_8U, Scalar(0));
         vector<Vec4i> lines;
 
         HoughLinesP(contours,lines,houghPDeltaRho,houghPDeltaTheta,houghPMinVote, houghPMinLineLength, houghPGap);
-        drawDetectedLines(houghP, lines);
+        drawDetectedLines(result, lines);
 
 
         if(showSteps) {
             namedWindow( "Input", CV_WINDOW_AUTOSIZE );
-            imshow( "Input", image );
[…]

### 2016-06-12 Headers/LineDetection.hpp
Tweaked edgedetection parameters

@@ -23,17 +23,18 @@ public:
     Mat getLinesFromImage(Mat &image, bool showSteps) {
 
 
+        // Filter hough detected lines by angle
+        int minAngle = 20;
+        int maxAngle = 80;
 
         // parameters
-        int cannyLowerThreshold = 28;
+        int cannyLowerThreshold = 40; // 40
         int cannyUpperThreshold = cannyLowerThreshold * 2;
-        int threshThreshold = 240;
-        int threshMaxval = 255;
-        int houghPMinLineLength = 60;
-        int houghPGap = 40;
-        int houghPMinVote = 5;
+        int houghPMinLineLength = 7;
+        int houghPGap = 10;
+        int houghPMinVote = 7;
         double houghPDeltaRho = 1;
-        double houghPDeltaTheta = PI/180;
+        double houghPDeltaTheta = PI / 180;
 
         // set the ROI for the image
         Rect roi(0, image.cols / 4, image.cols - 1, image.rows - image.cols / 4);
@@ -42,62 +43,59 @@ public:
 
         Mat tmp;
         Size size;
-        size.width  = 90;
-        size.height = 15;
+        size.width = 100; // 100
+        size.height = 30; // 20
         Mat kernel = getStructuringElement(MORPH_ELLIPSE, size);
-        //erode verticaal
-        erode(imgROI,tmp,kernel);
+        erode(imgROI, tmp, kernel);
 
-        size.width  = 60;
-        size.height = 10;
+        size.width = 85; // 85
+        size.height = 25; // 18
         kernel = getStructuringElement(MORPH_ELLIPSE, size);
-        dilate(tmp,tmp,kernel);
+        dilate(tmp, tmp, kernel);
 
 
         // Canny algorithm
         Mat contours;
-        Mat result;
+        Mat result(imgROI.size(), CV_8U, Scalar(0));
         Canny(tmp, contours, cannyLowerThreshold, cannyUpperThreshold, 3);
         Mat contoursInv;
-        //threshold(contours, contoursInv, threshThreshold, threshMaxval, THRESH_BINARY_INV);
         contours.copyTo(result);
 
         // Do probabilistic Hough
-        Mat houghP(imgROI.size(), CV_8U, Scalar(0));
         vector<Vec4i> lines;
 
[…]

### 2016-06-12 Headers/svm.hpp
Logging on parameter iterations for comparison

@@ -32,12 +32,14 @@ public:
 		svm = SVM::create();
 	    svm->setType(SVM::C_SVC);
 	    svm->setKernel(SVM::RBF);
+	    // svm->setKernel(SVM::LINEAR);
 	    svm->setTermCriteria(TermCriteria(TermCriteria::MAX_ITER, 100, 1e-6));
 	    Ptr<TrainData> trainData_ptr = TrainData::create(trainingsMat, ROW_SAMPLE , labelsMat);
 		cout << "training the SVM... "<<(trainAuto? "(automatically, this could take a while...)" : "") << endl;
 	    if(trainAuto){
 		    /* Automatic training */
-		    svm->trainAuto(trainData_ptr,10,SVM::getDefaultGrid(SVM::C),SVM::getDefaultGrid(SVM::GAMMA),ParamGrid(),ParamGrid(),ParamGrid(),ParamGrid(),true);
+		    ParamGrid emptyGrid(0,0,0);
+		    svm->trainAuto(trainData_ptr,10,SVM::getDefaultGrid(SVM::C),SVM::getDefaultGrid(SVM::GAMMA),emptyGrid,emptyGrid,emptyGrid,emptyGrid,true);
 		}else{
 		    /* Hardcoded params */
 		    svm->setC(io::C);

### 2016-06-12 trainingsdata.cpp

@@ -199,7 +199,7 @@ void parameterIterationTraining(bool relabel=true){
     int frameInterval = 10;
     // int frameStopIdx = io::FRAME_MAX_IDX;
     int frameStopIdx = 50;
-	bool trainAuto = false; // Whether or not to use automatic training for SVM
+	bool trainAuto = true; // Whether or not to use automatic training for SVM
 	bool includeMarksVals[] = {false,true};
 	bool useColorVals[] = {false,true,true};
 	bool useLBPVals[] = {true,false,true};
@@ -266,14 +266,30 @@ void parameterIterationTraining(bool relabel=true){
 	string fnFrameOutput = io::dirOutputLogging+"output_"+io::currentDateTime()+".csv";
 	csv.open(fnFrameOutput.c_str(),ios::out);
 
-	for(int i=0; i<nDatasets-1; i++){
-		for(int j=i+1;j<nDatasets;j++){
-			set<int> tS;
-			tS.insert(datasets[i]);
-			tS.insert(datasets[j]);
-			trainingsSets.push_back(tS);
+	bool useThreeTrainingsSets = false;
+	if(useThreeTrainingsSets){
+		for(int i=0; i<nDatasets-2; i++){
+			for(int j=i+1;j<nDatasets-1;j++){
+				for(int k=j+1;k<nDatasets;k++){
+					set<int> tS;
+					tS.insert(datasets[i]);
+					tS.insert(datasets[j]);
+					tS.insert(datasets[k]);
+					trainingsSets.push_back(tS);
+				}
+			}
+		}
+	}else{
+		for(int i=0; i<nDatasets-1; i++){
+			for(int j=i+1;j<nDatasets;j++){
+				set<int> tS;
+				tS.insert(datasets[i]);
+				tS.insert(datasets[j]);
+				trainingsSets.push_back(tS);
+			}
 		}
 	}
+
 	if(trainingsSets.empty() && nDatasets > 0){ // Single dataset
 		set<int> tS;
 		tS.insert(datasets[0]);

### 2016-06-13 Rapport/edgedetection.tex
Added some images for the paper and a new tex file for edgedetection

\chapter{Edge Detection}

As mentioned earlier, road detection is performed in squares of 32 px. When determining the exact transition from road to non-road, the best possible accuracy is 32 px. 
\npar --- nog verder aanvullen: edge detectie voor hogere nauwkeurigheid van de exacte overgang van weg naar niet-weg en voor fallback op wegdetectie bij slechte wegen, zoals kasseiwegen

\subsection{Removing details}

Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection. However, there are a lot of edges on the road itself, thinking of road markings and cobbles. The edges on the road, considerd as noise, and the edges of the road are not unambiguously separable. The width of the patterns that cause the edges on the road is rather small. Using morphological image processing operations, it is possible to remove these small patterns from the image.  
\npar
Eroding is

Kasseien: kassei zelf is het helderste (voorgrond),  krimpt --> voeg groeit, tot het één oppervlak vormt
Lijnen: Lijn is het helderste (voorgrond), krimpt --> verdwijnt

Nadien: alle donkere objecten groter geworden, compenseren via dilatie

\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{cobbles_in}
  \caption{Original frame\label{zebra_orig}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{cobbles_erode_out}
  \caption{Objects become bigger after eroding\label{zebrawitcanny}}
\end{subfigure}
\caption{Perform eroding and dilation to remove details.}
\end{figure}

\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{car_original}
  \caption{Original frame\label{zebra_orig}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{car_erode_bigger}
  \caption{Objects become bigger after eroding\label{zebrawitcanny}}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{car_dilate_smaller}
  \caption{Restore original sizes with dilation\label{zebrawitcanny}}
\end{subfigure}
\caption{Perform eroding and dilation to remove details.}
\end{figure}

\subsection{Canny Edge Detection}

### 2016-06-14 Rapport/edgedetection.tex
Explained eroding in edge-detection section of the paper

@@
 
-As mentioned earlier, road detection is performed in squares of 32 px. When determining the exact transition from road to non-road, the best possible accuracy is 32 px. 
-\npar --- nog verder aanvullen: edge detectie voor hogere nauwkeurigheid van de exacte overgang van weg naar niet-weg en voor fallback op wegdetectie bij slechte wegen, zoals kasseiwegen
-
+As mentioned earlier, road detection is performed in squares of 32 px.
+When determining the exact transition from road to non-road, the best possible accuracy is 32 px.
+To determine the precise transition from road to non-road, edge detection will be used.
+It can also be used as fallback when the road detection fails.
+\npar 
 \subsection{Removing details}
@@
 However, there are a lot of edges on the road itself, thinking of road markings and cobbles.
-The edges on the road, considerd as noise, and the edges of the road are not unambiguously separable.
+The edges on the road, considered as noise, and the edges of the road are not unambiguously separable.
 The width of the patterns that cause the edges on the road is rather small.
-Using morphological image processing operations, it is possible to remove these small patterns from the image.
+Using morphological image processing operations, it is possible to remove these small patterns.
 \npar
-Eroding is
-
-Kasseien: kassei zelf is het helderste (voorgrond),  krimpt --> voeg groeit, tot het één oppervlak vormt
-Lijnen: Lijn is het helderste (voorgrond), krimpt --> verdwijnt
+Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller.
+When the kernel is large enough, white lines can be filtered out completely, as seen in figure \ref{}.
+The same technique is used to smooth the surface of cobbled roads.
+A cobble stone consist of a bright center, surrounded by darker joints.
+By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints.
+With a large kernel, the joints will spread until they overlap.
+In figure \ref{}, a smoothed cobbled road is shown.
+\npar
+Unfortunately, the erode function will cause the darker objects to expand.
+The original size of the objects has been modified, as seen in \ref{}.
+This will falsify the eventual edge detection.
+In this example, the car is closer to the driver
 

### 2016-06-14 Rapport/edgedetection.tex
Paper: Edge Detection section readyµ

@@
 
-Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection.
+Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection \cite{canny}.
 However, there are a lot of edges on the road itself, thinking of road markings and cobbles.
@@
 Using morphological image processing operations, it is possible to remove these small patterns.
-\npar
-Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller.
-When the kernel is large enough, white lines can be filtered out completely, as seen in figure \ref{}.
-The same technique is used to smooth the surface of cobbled roads.
-A cobble stone consist of a bright center, surrounded by darker joints.
-By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints.
-With a large kernel, the joints will spread until they overlap.
-In figure \ref{}, a smoothed cobbled road is shown.
-\npar
-Unfortunately, the erode function will cause the darker objects to expand.
-The original size of the objects has been modified, as seen in \ref{}.
-This will falsify the eventual edge detection.
-In this example, the car is closer to the driver
-
-Nadien: alle donkere objecten groter geworden, compenseren via dilatie
-
+In figure \ref{cobbles_erode_out}, a smoothed cobbled road is shown.
 \begin{figure}[ht]
@@
   \includegraphics[width=.9\textwidth]{cobbles_in}
-  \caption{Original frame\label{zebra_orig}}
+  \caption{Original frame\label{cobbles_in}}
 \end{subfigure}%
@@
   \includegraphics[width=.9\textwidth]{cobbles_erode_out}
-  \caption{Objects become bigger after eroding\label{zebrawitcanny}}
+  \caption{Eroded cobbled road\label{cobbles_erode_out}}
 \end{subfigure}
-\caption{Perform eroding and dilation to remove details.}
+\caption{Perform eroding to remove the edges on the road.}
 \end{figure}
+
+\npar
+Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{erode_dilate}.
+Figure \ref{line_zoom_eroded} shows an example of a thinned white line.
+When the kernel is large enough, white lines can be filtered out completely.
+The same technique is used to smooth the surface of cobbled roads.
+A cobblestone consist of a bright center, surrounded by darker joints.
+By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints.
+With a large kernel, the joints will spread until they overlap.
+The best results were achieved with an ellipsoidal kernel of 100 x 30 px.
+
+\begin{figure}[ht]
+	\begin{minipage}[t]{.45\textwidth}
+		\centering
+		\begin{subfigure}[t]{.5\textwidth}
+		  \centering
+		  \includegraphics[width=.9\textwidth]{cobbles_zoom_original}
+		  \caption{Original frame\label{cobbles_zoom_original}}
+		\end{subfigure}%
+		\begin{subfigure}[t]{.5\textwidth}
+		  \centering
+		  \includegraphics[width=.9\textwidth]{cobbles_zoom_eroded}
+		  \caption{Bigger joints after eroding\label{cobbles_zoom_eroded}}
+		\end{subfigure}
+		\caption{The effect of eroding cobbled roads.}
+	\end{minipage}%
+	\hspace{0.05\textwidth}
+	\begin{minipage}[t]{.45\textwidth}
+		\centering
+		\begin{subfigure}[t]{.5\textwidth}
+		  \centering
+		  \includegraphics[width=.9\textwidth]{line_zoom_original}
+		  \caption{Original frame\label{line_zoom_original}}
+		\end{subfigure}%
+		\begin{subfigure}[t]{.5\textwidth}
+		  \centering
+		  \includegraphics[width=.9\textwidth]{line_zoom_eroded}
+		  \caption{Smaller lines after eroding\label{line_zoom_eroded}}
+		\end{subfigure}
+		\caption{The effect of eroding road markings.}
+	\end{minipage}
+\end{figure}
+
+\npar
+Unfortunately, the erode function will cause the darker objects to expand.
+The original size of the objects has been modified, as seen in \ref{car_erode_bigger}.
+This will falsify the eventual edge detection.
+In this example, the observed car is reported to close.
+In order to restore the original measures, dilation is executed.
+This is the opposite morphology operator of eroding.
+The combination of eroding and dilation is called opening.
+When dilation is executed with the same kernel size of the erode function, the original size of the dark objects will decrease to their original size, as seen in \ref{car_dilate_smaller}.
+Note that, after eroding, some details of the shape of the objects are lost.
+In order to provide a safety margin, the dilate function is executed with a slightly smaller kernel.
 
@@
   \includegraphics[width=.9\textwidth]{car_original}
-  \caption{Original frame\label{zebra_orig}}
+  \caption{Original frame\label{car_original}}
 \end{subfigure}%
@@
   \includegraphics[width=.9\textwidth]{car_erode_bigger}
-  \caption{Objects become bigger after eroding\label{zebrawitcanny}}
+  \caption{Dark objects are bigger after eroding\label{car_erode_bigger}}
 \end{subfigure}
@@
   \includegraphics[width=.9\textwidth]{car_dilate_smaller}
-  \caption{Restore original sizes with dilation\label{zebrawitcanny}}
+  \caption{Restore original sizes with dilation\label{car_dilate_smaller}}
 \end{subfigure}
@@
 \subsection{Canny Edge Detection}
+
+To detect the edges of the road, the Canny Edge Detection algorithm is used. Figure \ref{canny_edges} shows an example.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=.5\textwidth]{canny_edges}
+	\caption{Canny edge detection\label{canny_edges}}
+\end{figure}
+
+\npar
+This algorithm requires a minimum and a maximum threshold value, which were determined experimentally. 
+The best results were achieved with a maximum threshold that is twice the minimum threshold.
+The minimum threshold is determined by the required minimum sensitivity.
+When the algorithm is configured too sensitive, some imperfections on the road will be detected as edges.
+On the other hand, a certain level of sensitivity is necessary to detect less clear road edges.
+For example, a transition from road to dirt has a very small distinction of color and brightness, as seen in \ref{transition_road_dirt_eroded}.
+The minimum threshold should be high enough to detect this transition as edge.
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{transition_road_dirt_original}
+  \caption{Original frame\label{transition_road_dirt_original}}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{transition_road_dirt_eroded}
+  \caption{Eroded road\label{transition_road_dirt_eroded}}
+\end{subfigure}
+\caption{Unclear transition from road to non-road.}
+\end{figure}

### 2016-06-14 Administratie/report/paper.tex
report SVM + edge

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{url}   
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Speed estimation using SVM classification and edge detection}

% a short form should be given in case it is too long for the running head
\titlerunning{Speed estimation using SVM classification and edge detection}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Mathias Dierickx\and Tim Ranson\and Maarten Tindemans\and Bjorn Vandenbussche}
%
\authorrunning{Speed estimation using SVM classification and *blablabla* line detection}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Ghent University, Faculty of Engineering and Architecture\\
Valentin Vaerwyckweg 1, 9000, Ghent}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\maketitle

\begin{abstract}
This is the abstract: TODO
\keywords{SVM classification, Local Binary Patterns, line detection}
\end{abstract}

\section{Introduction}
TODO

\section{Background}
\subsection{Texture analysis using Local Binary Patterns}
Local Binary Patterns (LBP) \cite{Huang2011} is a non-parametric descriptor whose aim is to efficiently summarize the local structures of images. As a non-parametric method, LBP summarizes local structures of images efficiently by comparing each pixel with its neighboring pixels. The most important properties of LBP are its tolerance regarding monotonic illumination changes and its computational simplicity. LBP was originally proposed for texture analysis \cite{Ojala1996}, and has proved a simple yet powerful approach to describe local structures.

The original LBP operator labels the pixels of an image with
decimal numbers, called Local Binary Patterns or LBP codes, which encode the local structure around each pixel (see figure \ref{fig:lbp}). Each pixel is then compared with its eight neighbors in a 3x3 neighborhood by subtracting the center pixel value. The resulting strictly negative values are encoded with 0 and the others with 1. A binary number is obtained by concatenating all these binary codes in a clockwise direction starting from the top-left one and its corresponding decimal value is used for labeling. The derived binary numbers are referred to as Local Binary Patterns or LBP codes.

One limitation of the basic LBP operator is that its small 3x3 neighborhood cannot capture dominant features with large scale structures. To deal with the texture at different scales, the operator was later generalized to use neighborhoods of different sizes \cite{Ojala}. The histogram of LBP labels can then be exploited as a texture descriptor.

\begin{figure}[h]
\centering
\includegraphics[height=3cm]{fig/LBP.png}
\caption{An example of the basic LBP operator}
\label{fig:lbp}
\end{figure}

\subsection{SVM classification}
Support Vector Machines (SVM's) \cite{Boswell2002} are a learning method used for binary classification. The basic idea is to find a hyperplane which separates the $d$-dimensional data perfectly into its two classes.

TODO: Tim
\subsection{Precision, recall, F-scores}

TODO: Bjorn

\subsection{Line detection}
TODO: Mathias

\section{Method}
In order to determine the maximum speed of a vehicle two image processing techniques are used: SVM classification and edge detection. We first try to rely on edge detection to find the contours of the lanes where driving is permitted (i.e. where no obstacles are found). We estimate the reliability of the edge detection and incorporate road detection using the SVM classifier when needed.

For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++.

\subsection{Parameters for SVM classification}\label{parameters-classification}
The goal of this classifier is to determine what pixels are part of the road and which are not. We will describe multiple tunable parameters in order to get optimal classification for road detection. These parameters determine both the length and values of the feature vectors, which are then fed to the SVM classifier. We compare classification results varying these parameters for the feature vector in section \ref{results-svm-classification}.

\subsubsection{Block size}
Each frame is split into multiple blocks of predetermined size. We chose to vary the block sizes between 8x8px, 16x16px and 32x32px. On the one hand, larger block sizes would too easily contain multiple structures of different classes, resulting in poorer classification. For example a block on the edge of the road containing part of the street and grass. On the other hand, smaller block sizes would omit too many details to be descriptive enough to use for classification.

The block size affects the level of detail taken into account while calculating the feature vector, as the values for the feature vector operate within one block size.

TODO: insert photos block sizes

\subsubsection{LBP}
With LBP we try to characterize the structure of roads within a block. The histogram of LBP values within that block can be added to the feature vector in order to improve classification. These histograms are calculated per channel (e.g. 3 channels for RGB) and are added consecutively to the feature vector.

\subsubsection{Color}
Because working with LBP values only doesn't take color intensity into account \cite{Pietikainen2002}, histogram values of color intensity per block and channel are added to the feature vector.

\subsection{Kernels for SVM classification}\label{methods-kernels}
Judging from the OpenCV documentation\footnote{http://docs.opencv.org/3.1.0/} there are two SVM kernels fit for a two-class classification problem.

TODO: Tim

\subsection{Road marks}
The training dataset was labeled considering blocks containing road marks once as part of the road class and once as part of the non-road class. Initially road marks were labeled as non-road, in order to get a good classification of road (i.e. different types of asphalt). As these road marks aren't really considered an obstacle for a car driver, ideally they should be classified as road. This makes up for a more difficult classification problem. To quantify this effect we apply classification both with and without road marks.

\subsection{Edge detection}

\subsection{Removing details}

Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection \cite{canny}. However, there are a lot of edges on the road itself, thinking of road markings and cobbles. The edges on the road, considered as noise, and the edges of the road are not unambiguously separable. The width of the patterns that cause the edges on the road is rather small. Using morphological image processing operations, it is possible to remove these small patterns.  In figure \ref{cobbles_erode_out}, a smoothed cobbled road is shown.
\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_in.png}
  \caption{Original frame\label{cobbles_in}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_erode_out.png}
  \caption{Eroded cobbled road\label{cobbles_erode_out}}
\end{subfigure}
\caption{Perform eroding to remove the edges on the road.}
\end{figure}

Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{erode_dilate}. Figure \ref{line_zoom_eroded} shows an example of a thinned white line. When the kernel is large enough, white lines can be filtered out completely. The same technique is used to smooth the surface of cobbled roads. A cobblestone consist of a bright center, surrounded by darker joints. By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints. With a large kernel, the joints will spread until they overlap. The best results were achieved with an ellipsoidal kernel of 100 x 30 px.

\begin{figure}[ht]
	\begin{minipage}[t]{.45\textwidth}
		\centering
		\begin{subfigure}[t]{.5\textwidth}
		  \centering
		  \includegraphics[width=.9\textwidth]{fig/cobbles_zoom_original.png}
		  \caption{Original frame\label{cobbles_zoom_original}}
		\end{subfigure}%
		\begin{subfigure}[t]{.5\textwidth}
		  \centering
		  \includegraphics[width=.9\textwidth]{fig/cobbles_zoom_eroded.png}
		  \caption{Bigger joints after eroding\label{cobbles_zoom_eroded}}
		\end{subfigure}
		\caption{The effect of eroding cobbled roads.}
	\end{minipage}%
	\hspace{0.05\textwidth}
	\begin{minipage}[t]{.45\textwidth}
		\centering
		\begin{subfigure}[t]{.5\textwidth}
		  \centering
		  \includegraphics[width=.9\textwidth]{fig/line_zoom_original.png}
		  \caption{Original frame\label{line_zoom_original}}
		\end{subfigure}%
		\begin{subfigure}[t]{.5\textwidth}
		  \centering
		  \includegraphics[width=.9\textwidth]{fig/line_zoom_eroded.png}
		  \caption{Smaller lines after eroding\label{line_zoom_eroded}}
		\end{subfigure}
		\caption{The effect of eroding road markings.}
	\end{minipage}
\end{figure}

Unfortunately, the erode function will cause the darker objects to expand. The original size of the objects has been modified, as seen in \ref{car_erode_bigger}. This will falsify the eventual edge detection. In this example, the observed car is reported to close. In order to restore the original measures, dilation is executed. This is the opposite morphology operator of eroding. The combination of eroding and dilation is called opening. When dilation is executed with the same kernel size of the erode function, the original size of the dark objects will decrease to their original size, as seen in \ref{car_dilate_smaller}. Note that, after eroding, some details of the shape of the objects are lost. In order to provide a safety margin, the dilate function is executed with a slightly smaller kernel.

\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/car_original.png}
  \caption{Original frame\label{car_original}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/car_erode_bigger.png}
  \caption{Dark objects are bigger after eroding\label{car_erode_bigger}}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/car_dilate_smaller.png}
  \caption{Restore original sizes with dilation\label{car_dilate_smaller}}
\end{subfigure}
\caption{Perform eroding and dilation to remove details.}
\end{figure}

\subsection{Canny Edge Detection}

To detect the edges of the road, the Canny Edge Detection algorithm is used. Figure \ref{canny_edges} shows an example.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.5\textwidth]{fig/canny_edges.png}
	\caption{Canny edge detection\label{canny_edges}}
\end{figure}

This algorithm requires a minimum and a maximum threshold value, which were determined experimentally. 
The best results were achieved with a maximum threshold that is twice the minimum threshold.
The minimum threshold is determined by the required minimum sensitivity. When the algorithm is configured too sensitive, some imperfections on the road will be detected as edges. On the other hand, a certain level of sensitivity is necessary to detect less clear road edges. For example, a transition from road to dirt has a very small distinction of color and brightness, as seen in \ref{transition_road_dirt_eroded}. The minimum threshold should be high enough to detect this transition as edge.

\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/transition_road_dirt_original.png}
  \caption{Original frame\label{transition_road_dirt_original}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/transition_road_dirt_eroded.png}
  \caption{Eroded road\label{transition_road_dirt_eroded}}
\end{subfigure}
\caption{Unclear transition from road to non-road.}
\end{figure}

\subsection{Combination of classification and edge detection}

TODO: Maarten

\section{Results}

\subsection{SVM classification}\label{results-svm-classification}
We iterated through several parameter values (as discussed in \ref{parameters-classification}) to find optimal classification for road detection. For each combination of parameters we train and test on different combinations of datasets. For each combination we calculate the F-scores. These scores are used to compare the classification results (higher F-scores indicate better classification).

Of the four datasets, different combinations of training sets were selected (each time shown as x-axis on the graphs). The trained classifier was then tested on the remaining datasets.

\subsubsection{Kernels}
As previously mentioned in section \ref{methods-kernels}, there are two SVM kernels we can use for this classification problem. On average, we tend to get a nearly 11\% increase in F-scores using RBF kernel instead of linear kernel. It is safe to say the RBF kernel outperforms the linear kernel for this classification problem in all of our test scenarios.

\subsubsection{Block size}
In general, block sizes 16x16px and 32x32px tend to show the better results. As an example we present the F-scores for LBP and color with road marks in figure \ref{fig:lbp-col-wi-rm}. Other combinations follow the same trend. For the goal of this application, a smaller block size will be more practical (i.e. better accuracy). This is why prefer the block size of 16x16px.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig/RBF_LBP_Col_wi_RM.png}
\caption{F-scores for LBP and color with road marks using RBF kernel}
\label{fig:lbp-col-wi-rm}
\end{figure}

\subsubsection{LBP}
The use of LBP is crucial to our implementation in order to get a viable classifier for road detection. If we compare results (block size 16x16px) of feature vectors on LBP only to the combination of LBP and color, we get different results depending on whether or not we include road marks.

If we do not include road marks in the training data, we get slightly better results when combining LBP and color for the feature vectors (as shown in figure \ref{fig:16_wo_roadmarks}).
If on the other hand we do include these road marks, using only LBP shows equal or better results than the combination of LBP and color (as shown in figure \ref{fig:16_wi_roadmarks}).

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/16_wo_roadmarks.png}
\caption{Comparision of F-scores for 16x16px block size without road marks using RBF kernel }
\label{fig:16_wo_roadmarks}
\end{figure}

A possible explanation for this phenomenon is that when only asphalt is considered to be road, color is a much more consistent factor to weigh in to the classification. However when including road marks as part of the road, color intensity is no longer an added value and even tends to mislead the classifier.

Overall we can conclude that using only LBP with road marks shows the optimal results for our classification problem.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/16_wi_roadmarks.png}
\caption{Comparision of F-scores for 16x16px block size with road marks using RBF kernel }
\label{fig:16_wi_roadmarks}
\end{figure}
\newpage
\noindent In retrospect, the optimal parameters for our classification problem are:
\begin{itemize}
\item RBF kernel
\item Block size 16x16px
\item Using only LBP histogram values for the feature vectors
\item Including road marks as part of the road  
\end{itemize}

\subsection{Edge detection}

\section{Conclusion}
\newpage
\bibliography{Computervisie}{}
\bibliographystyle{plain}

\end{document}

### 2016-06-14 Administratie/report/paper.tex
update report
title change

@@
 % first the title is needed
-\title{Speed estimation using SVM classification and edge detection}
+\title{Speed estimation for vehicles using \\SVM classification and edge detection}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{Speed estimation using SVM classification and edge detection}
+\titlerunning{Speed estimation for vehicles using SVM classification and edge detection}
 
@@
 %
-\authorrunning{Speed estimation using SVM classification and *blablabla* line detection}
+\authorrunning{Speed estimation for vehicles using SVM classification and edge detection}
 % (feature abused for this document to repeat the title also on left hand pages)
@@
 
-For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++.
+For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either or test the solution.
+
+\textbf{TODO}: kort over training van classificatie  (TIM)
 
@@
 The block size affects the level of detail taken into account while calculating the feature vector, as the values for the feature vector operate within one block size.
+Some examples of different block sizes can be found in figure \ref{fig:methods-block-sizes}.
 
-TODO: insert photos block sizes
+TODO
 
@@
 To quantify this effect we apply classification both with and without road marks.
-
-\subsection{Edge detection}
 

### 2016-06-14 Administratie/report/paper.tex
update

@@
 \end{figure}
-\newpage
+
+Considering the results of the different scenarios, we can conclude that training on dataset four results in poorer classification overall.
+This is because another type of road is used in this dataset, conflicting with the more common road type present in the other datasets.
+This distorts the classifier increases misclassifications.
+
 \noindent In retrospect, the optimal parameters for our classification problem are:

### 2016-06-14 Headers/io.hpp
Report SVM blockgrid examples

@@ -32,17 +32,16 @@ namespace io {
     const int frameInterval = 10;
     const int frameStopIdx = 50;
     const int innerMargin = 64;
-    const int blkSize = 32; // 8,16,32
+    const int blkSize = 16; // 8,16,32
     const int lbpRadius = 1;
-    const bool includeMarks = false;
+    const bool includeMarks = true;
     const bool trainAuto = false;
     const bool useLBP = true; // Make sure to use either color, LBP or both
-    const bool useColor = true;
+    const bool useColor = false;
 
     /** SVM parameters for when not training automatically **/
     const double C = 0.1;
     const double gamma = 0.00001;
-    // const double gamma = 0.00015;
 
 	/** Shown image properties and helper values **/
 	int imX=50,imY=50

### 2016-06-14 Headers/svm.hpp

@@ -125,18 +125,18 @@ void my_svm::calculateScores(Mat& extLabels, Mat& extTrainingsMat) {
 		        confusion.at<int>(0,1)++;
 		    }
 		}
-		accuracy = ((double)(confusion.at<int>(0,0)+confusion.at<int>(1,1)))/((double)extTrainingsMat.rows)*100;
-		precision = ((double)confusion.at<int>(0,0)/(confusion.at<int>(0,0)+confusion.at<int>(1,0))*100);
-		recall = ((double)confusion.at<int>(0,0)/(confusion.at<int>(0,0)+confusion.at<int>(0,1)))*100;
-		true_negative = ((double)confusion.at<int>(1,1)/(confusion.at<int>(1,1)+confusion.at<int>(1,0)))*100;	
+		accuracy = ((double)(confusion.at<int>(0,0)+confusion.at<int>(1,1)))/((double)extTrainingsMat.rows);
+		precision = (double)confusion.at<int>(0,0)/(confusion.at<int>(0,0)+confusion.at<int>(1,0));
+		recall = (double)confusion.at<int>(0,0)/(confusion.at<int>(0,0)+confusion.at<int>(0,1));
+		true_negative = (double)confusion.at<int>(1,1)/(confusion.at<int>(1,1)+confusion.at<int>(1,0));	
 		F = 2*(precision*recall)/(precision+recall);
 	}
     cout<<"Confusion matrix: "<<endl<<confusion<<endl;
-    cout<<"Precision: "<<precision<<"%"<<endl;
-    cout<<"Recall: "<<recall<<"%"<<endl;
-    cout<<"Accuracy: "<<accuracy<<"%"<<endl;
-    cout<<"True Negative Rate: "<<true_negative<<"%"<<endl;    
-    cout<<"F: "<<F<<"%"<<endl;
+    cout<<"Precision: "<<precision*100<<"%"<<endl;
+    cout<<"Recall: "<<recall*100<<"%"<<endl;
+    cout<<"Accuracy: "<<accuracy*100<<"%"<<endl;
+    cout<<"True Negative Rate: "<<true_negative*100<<"%"<<endl;    
+    cout<<"F: "<<F*100<<"%"<<endl;
 }
 
 /**
@@ -177,6 +177,9 @@ void my_svm::test(int min = 1, int max = 4, int skip = 5, int number = 5) {
                 );
             }
             imshow("output", image); // show it to the user
[…]

### 2016-06-14 svm.cpp

@@ -1,16 +1,18 @@
 #include "Headers/lbpfeaturevector.hpp"
 #include "Headers/io.hpp"
 #include "Headers/svm.hpp"
+#include <set>
 
 using namespace std;
 using namespace cv;
 
 int main(int argc, char** argv){
 
-    vector<short> trainingDatasets,testDatasets;
-    trainingDatasets.push_back(1);
-    testDatasets.push_back(1);
     Mat initLabels,initTraining,testLabels, testTraining;
+
+    set<int> tS;
+    tS.insert(io::datasets[0]);
+    tS.insert(io::datasets[2]);
 
     string frameName;
     char buffer[30];
@@ -18,19 +20,19 @@ int main(int argc, char** argv){
     int dataset = io::datasets[dSIdx];
     for(int fIdx=0;fIdx<io::frameStopIdx;fIdx+=io::frameInterval){
         io::buildFrameName(buffer,frameName,dataset,fIdx,io::innerMargin,io::blkSize,io::includeMarks);
-        io::readTrainingsdataOutput(frameName,initLabels,initTraining,io::useLBP,io::useColor);
-        io::readTrainingsdataOutput(frameName,testLabels,testTraining,io::useLBP,io::useColor);
+        bool isTrainingSet = tS.find(dataset)!=tS.end();
+        if(isTrainingSet)
+            io::readTrainingsdataOutput(frameName,initLabels,initTraining,io::useLBP,io::useColor);
+        else
+            io::readTrainingsdataOutput(frameName,testLabels,testTraining,io::useLBP,io::useColor);
     }
     }
-
-    // io::readTrainingsdata(trainingDatasets,initLabels,initTraining);
-    // io::readTrainingsdata(testDatasets,testLabels,testTraining);
 
     my_svm svm(initLabels,initTraining,io::trainAuto);
     svm.calculateScores(testLabels,testTraining);
 
     //use for visual testing
-    svm.test(1,4,10,5);
+    svm.test(1,1,10,5);
 
 	return 0;
 }

### 2016-06-14 trainingsdata.cpp

@@ -266,8 +266,19 @@ void parameterIterationTraining(bool relabel=true){
 	string fnFrameOutput = io::dirOutputLogging+"output_"+io::currentDateTime()+".csv";
 	csv.open(fnFrameOutput.c_str(),ios::out);
 
-	bool useThreeTrainingsSets = false;
-	if(useThreeTrainingsSets){
+	bool useCoupleTrainingsSets = true;
+	bool useTripletTrainingsSets = true;
+	if(useCoupleTrainingsSets){
+		for(int i=0; i<nDatasets-1; i++){
+			for(int j=i+1;j<nDatasets;j++){
+				set<int> tS;
+				tS.insert(datasets[i]);
+				tS.insert(datasets[j]);
+				trainingsSets.push_back(tS);
+			}
+		}
+	}
+	if(useTripletTrainingsSets){
 		for(int i=0; i<nDatasets-2; i++){
 			for(int j=i+1;j<nDatasets-1;j++){
 				for(int k=j+1;k<nDatasets;k++){
@@ -277,15 +288,6 @@ void parameterIterationTraining(bool relabel=true){
 					tS.insert(datasets[k]);
 					trainingsSets.push_back(tS);
 				}
-			}
-		}
-	}else{
-		for(int i=0; i<nDatasets-1; i++){
-			for(int j=i+1;j<nDatasets;j++){
-				set<int> tS;
-				tS.insert(datasets[i]);
-				tS.insert(datasets[j]);
-				trainingsSets.push_back(tS);
 			}
 		}
 	}

### 2016-06-15 Administratie/report/paper.tex
Report SVM and training

@@
 Support Vector Machines (SVM's) \cite{Boswell2002} are a learning method used for binary classification.
-The basic idea is to find a hyperplane which separates the $d$-dimensional data perfectly into its two classes.
+The idea is to find a hyperplane which separates the $d$-dimensional data perfectly into its two classes.
+However, since example data is often not linearly separable, SVM's introduce the notion of a "kernel induced feature space" which casts the data into a higher dimensional space where the data is separable.
+Typically, casting into such a space would cause problems computationally, and with overfitting.
+The key insight used in SVM's is that the higher-dimensional space doesn't need to be dealt with directly, which eliminates the above concerns.
 
-TODO: Tim
+Basically, we want the hyperplane that maximizes the geometric distance to the closest data points (as shown in figure \ref{fig:hyperplane}).
+
+\begin{figure}[h]
+\centering
+\includegraphics[height=7cm]{fig/SVM_hyperplane.png}
+\caption{Choosing the hyperplane that maximizes the margin.}
+\label{fig:hyperplane}
+\end{figure}
+
+Only the closest data points contribute to deciding the optimal hyperplane, which is why these are called the \textit{support vectors}.
+They are the only ones needed in defining (and finding) the optimal hyperplane.
+
+The equation for the hyperplane is subject to a constraint concerning the parameter $C$. 
+One can think of $C$ as a tunable parameter: higher $C$ corresponds to more importance on classifying all the training data correctly, lower $C$ results in a more "flexible" hyperplane that tries to minimize the margin error for each example.
+
+\paragraph{Kernels} The original equation for the hyperplane to linearly separate the data contains a linear kernel. \cite{Hsu2016} points out when the use of a simple linear kernel is appropriate:
+if the number of features is large, one may not need to map data to a higher dimensional space.
+That is, the nonlinear mapping does not improve the performance.
+Using the linear kernel is good enough, and one only searches for the parameter $C$.
+
+We mentioned that SVM's can use a "kernel induced feature space" which casts the data into a higher dimensional space where the data is separable, in case the data is not linearly separable in its original space.
+Pending on the type of data and application of the SVM, there are several types of kernels which could be used.
+The Radial Basis Function (RBF) kernel is a often a reasonable first choice, because it can handle the case when the relation between class labels and attributes is non-linear, while only adding one more parameter $\gamma$ to search for.
+
+\paragraph{Parameters} There are two parameters for an RBF kernel: $C$ and $\gamma$.
+It is not known beforehand which $C$ and $\gamma$ are best for a given problem; consequently some kind of model selection (parameter search) must be done.
+The goal is to identify good ($C$, $\gamma$) so that the classifier can accurately predict unknown data (i.e.
+testing data).
+
+One method for comparing results for a set of parameter values is cross-validation.
+In $v$-fold cross-validation, we first divide the training set into $v$ subsets of equal size.
+Sequentially one subset is tested using the classifier trained on the remaining $v − 1$ subsets.
+Thus, each instance of the whole training set is predicted once so the cross-validation accuracy is the percentage of data which are correctly classified.
+
 \subsection{Precision, recall, F-scores}
@@
 
-For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either or test the solution.
-
-\textbf{TODO}: kort over training van classificatie  (TIM)
+For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either train or test the solution.
+To train out SVM classifier, we used a function of the OpenCV SVM library which selects the optimal parameters using cross-validation.
 
@@
 
-TODO
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.49\textwidth}
+  \centering
+  \includegraphics[width=1\textwidth]{fig/svm_prediction_16_lbp_marks.png}
+  \caption{Block size 16x16px}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+  \centering
+  \includegraphics[width=1\textwidth]{fig/svm_prediction_32_lbp_marks.png}
+  \caption{Block size 32x32px}
+\end{subfigure}
+\caption{SVM prediction using only LBP values and including road marks. Green blocks were predicted to be road, red to be non-road.}
+\label{fig:methods-block-sizes}
+\end{figure}
 
@@
 Judging from the OpenCV documentation\footnote{http://docs.opencv.org/3.1.0/} there are two SVM kernels fit for a two-class classification problem.
+A linear kernel is faster, but the RBF kernel will outperform the linear kernel in case the data is not linearly separable.
+We tested and compared the results using each kernel to decide which is best fit for our classification problem.
 
-TODO: Tim
+\subsection{Training the SVM classifier}
+For each combination of the previously mentioned parameters, the SVM classifier must be trained.
+For a training set (a subset of the datasets available) each frame is divided into blocks.
+Each block must be manually labelled with the expected value for classification.
+For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset (i=0,10,...,40).
+Figure \ref{fig:frame-training} shows a visualization of pre-labeled blocks in a frame.
+The classifier is trained using the label and calculated feature vector for each block in each frame of the training set.
+
+\begin{figure}[ht]
+\centering
+  \centering
+  \includegraphics[width=.7\textwidth]{fig/svm_training_32.png}
+\caption{Visualization of a pre-labelled frame of a training set. Block size 32x32px, excluding road marks}
+\label{fig:frame-training}
+\end{figure}
 
 \subsection{Road marks}
-The training dataset was labeled considering blocks containing road marks once as part of the road class and once as part of the non-road class.
-Initially road marks were labeled as non-road, in order to get a good classification of road (i.e.
+The training dataset was labelled considering blocks containing road marks once as part of the road class and once as part of the non-road class.
+Initially road marks were labelled as non-road, in order to get a good classification of road (i.e.
 different types of asphalt).
@@
 
+\textbf{TODO} check if list not f***ed up.\\
 \noindent In retrospect, the optimal parameters for our classification problem are:

### 2016-06-15 trainingsdata.cpp

@@ -199,7 +199,7 @@ void parameterIterationTraining(bool relabel=true){
     int frameInterval = 10;
     // int frameStopIdx = io::FRAME_MAX_IDX;
     int frameStopIdx = 50;
-	bool trainAuto = true; // Whether or not to use automatic training for SVM
+	bool trainAuto = false; // Whether or not to use automatic training for SVM
 	bool includeMarksVals[] = {false,true};
 	bool useColorVals[] = {false,true,true};
 	bool useLBPVals[] = {true,false,true};

### 2016-06-15 Administratie/report/paper.tex
Report update: combination road classification, edge detection

@@
 
-TODO: Maarten
+To combine road classification and edge detection we should first look at the performance of these two methods separately.
+Therefore, the evaluation was run on each of these two methods.
+The result is shown in figure \ref{distance_result}.
+The first bar of the histogram shows for every dataset the performance of road classification.
+Because of a crash in the first dataset there can only be travelled 317m.
+The second bar shows for each dataset the distance travelled only relying on edge detection.
+There are no crashes in the four datasets so the results are better.
+
+A first and very safe way to combine these two methods is the following.
+A region is considered road only if it's identified as road by the road classification and there is no edge.
+The performance of this combination is shown in the third bar.
+As said this is very safe, therefore the travelled distance is lower than only applying edge detection.
+
+An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
+A region is considered road if there are no edges.
+When there are edges, road classification is used to decide whether it's road.
+This strategy gave us the best results and are shown as the fourth bar.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\textwidth]{fig/distance_result.png}
+	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
+\end{figure}
+
+This last combination based on edge detection is able to cope with zebra crossings.
+This is shown in figure \ref{zebra}.
+In the first one the road markings are small because the car is far enough.
+Edge detection is therefore able to filter out these markings.
+In the next frame the markings are closer to the care and are too large to be filtered out.
+Because the road classification suggests the markings are road, they are considered road.
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
+  \caption{Zebra crossing filtered out by edge detection.}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
+  \caption{Zebra crossing classified as road.}
+\end{subfigure}
+\caption{Zebra crossing\label{zebra}}
+\end{figure}
+
+The combination of road classification and edge detection starts with removing outliers from the road classification.
+The goal of this is to eliminate the effect an outlier would have when it's close to the car.
+A non-road block is considered an outlier when it has less than 4 non-road block neighbours.
+This is repeated until we become a stable condition without outliers.
+Next, the edges are checked if they aren't on the road.
+This is done by checking that an edge is on a non-road block of the road classification.
+For safety reasons there is an additional check.
+The above, left and right blocks of an edge should also be considered as non-road.
+
+When the edge is not classified on the road, it is added to a collection.
+The intersections of the edges in the collection and the masks are computed.
+For each frame the masks give the projected trajectory of the car.
+The edges are visualised as blue lines on figure \ref{zebra}.
+The intersections are shown green.
+The closest intersection to the car is shown in red.
+At this point the maximum speed is derived from the intensity.
 

### 2016-06-15 Administratie/report/paper.tex
image allignment
reorganisation
Changed cite to paper about erode/dilate

@@
 
-Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{erode_dilate}.
+Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{sreedhar2012enhancement}.
 Figure \ref{line_zoom_eroded} shows an example of a thinned white line.
@@
 
-\subsection{Combination of classification and edge detection}
-
-To combine road classification and edge detection we should first look at the performance of these two methods separately.
-Therefore, the evaluation was run on each of these two methods.
-The result is shown in figure \ref{distance_result}.
-The first bar of the histogram shows for every dataset the performance of road classification.
-Because of a crash in the first dataset there can only be travelled 317m.
-The second bar shows for each dataset the distance travelled only relying on edge detection.
-There are no crashes in the four datasets so the results are better.
-
-A first and very safe way to combine these two methods is the following.
-A region is considered road only if it's identified as road by the road classification and there is no edge.
-The performance of this combination is shown in the third bar.
-As said this is very safe, therefore the travelled distance is lower than only applying edge detection.
-
-An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
-A region is considered road if there are no edges.
-When there are edges, road classification is used to decide whether it's road.
-This strategy gave us the best results and are shown as the fourth bar.
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=\textwidth]{fig/distance_result.png}
-	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
-\end{figure}
+\subsection{Zebra crossings}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}[t]{.5\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}[t]{.5\textwidth}
   \centering
@@
 
+\subsection{Combination of classification and edge detection}
+
 The combination of road classification and edge detection starts with removing outliers from the road classification.
@@
 This is repeated until we become a stable condition without outliers.
+
 Next, the edges are checked if they aren't on the road.
@@
 
-\subsection{Edge detection}
+\subsection{Driving results}
+To combine road classification and edge detection we should first look at the performance of these two methods separately.
+Therefore, the evaluation was run on each of these two methods.
+The result is shown in figure \ref{distance_result}.
+The first bar of the histogram shows for every dataset the performance of road classification.
+Because of a crash in the first dataset there can only be travelled 317m.
+The second bar shows for each dataset the distance travelled only relying on edge detection.
+There are no crashes in the four datasets so the results are better.
+
+A first and very safe way to combine these two methods is the following.
+A region is considered road only if it's identified as road by the road classification and there is no edge.
+The performance of this combination is shown in the third bar.
+As said this is very safe, therefore the travelled distance is lower than only applying edge detection.
+
+An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
+A region is considered road if there are no edges.
+When there are edges, road classification is used to decide whether it's road.
+This strategy gave us the best results and are shown as the fourth bar.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\textwidth]{fig/distance_result.png}
+	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
+\end{figure}
+
+This last combination based on edge detection is able to cope with zebra crossings.
+This is shown in figure \ref{zebra}.
+In the first one the road markings are small because the car is far enough.
+Edge detection is therefore able to filter out these markings.
+In the next frame the markings are closer to the care and are too large to be filtered out.
+Because the road classification suggests the markings are road, they are considered road.
 

### 2016-06-15 Administratie/report/paper.tex
update

@@
 
-Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{sreedhar2012enhancement}.
+Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{erode_dilate}.
 Figure \ref{line_zoom_eroded} shows an example of a thinned white line.
@@
 
-This last combination based on edge detection is able to cope with zebra crossings.
-This is shown in figure \ref{zebra}.
-In the first one the road markings are small because the car is far enough.
-Edge detection is therefore able to filter out these markings.
-In the next frame the markings are closer to the care and are too large to be filtered out.
-Because the road classification suggests the markings are road, they are considered road.
+\subsection{Combination of classification and edge detection}
 
-\begin{figure}[t]
-\centering
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
-  \caption{Zebra crossing filtered out by edge detection.}
-\end{subfigure}%
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
-  \caption{Zebra crossing classified as road.}
-\end{subfigure}
-\caption{Zebra crossing\label{zebra}}
-\end{figure}
-
-\subsection{Combination of classification and edge detection}
+In order to combine road classification and edge detection two methods were tested:
+\begin{enumerate}
+\item  A region is considered road only if it's identified as road by the road classification and there are no edges. 
+\item A region is considered road if there are no edges. When there are edges, road classification is used to decide whether it's road. 
+\end{enumerate}
 
@@
 A non-road block is considered an outlier when it has less than 4 non-road block neighbours.
-This is repeated until we become a stable condition without outliers.
+This is repeated until all outliers are removed.
 
@@
 
-When the edge is not classified on the road, it is added to a collection.
+When the edge is not classified as road, it is added to a collection.
 The intersections of the edges in the collection and the masks are computed.
 For each frame the masks give the projected trajectory of the car.
-The edges are visualised as blue lines on figure \ref{zebra}.
+The edges are visualised as blue lines on figure \ref{visualization}.
 The intersections are shown green.
@@
 At this point the maximum speed is derived from the intensity.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/frame00025.png}
+  \caption{Visualization of a processed frame showing both classification and edge detection}
+  \label{visualization}
+\end{figure}
 
@@
 
-An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
+An other way to make a combination, which still results in zero crashes is to focus on edge detection.
 A region is considered road if there are no edges.
@@
 
+\begin{figure}[t]
+\centering
+\begin{subfigure}[t]{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
+  \caption{Zebra crossing filtered out by edge detection.}
+\end{subfigure}%
+\begin{subfigure}[t]{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
+  \caption{Zebra crossing classified as road.}
+\end{subfigure}
+\caption{Zebra crossing\label{zebra}}
+\end{figure}
+
 \section{Conclusion}

### 2016-06-15 Administratie/report/paper.tex
Dataset output to separate folders

--- 2016-06-15 Administratie/report/paper.tex
@@
 \subsection{Combination of classification and edge detection}
+To combine road classification and edge detection we should first look at the performance of these two methods separately.
+Therefore, the evaluation was run on each of these two methods.
+The result is shown in figure \ref{distance_result}.
+The first bar of the histogram shows for every dataset the performance of road classification.
+Because of a crash in the first dataset there can only be travelled 317m.
+The second bar shows for each dataset the distance travelled only relying on edge detection.
+There are no crashes in the four datasets so the results are better.
+
+A first and very safe way to combine these two methods is the following.
+A region is considered road only if it's identified as road by the road classification and there is no edge.
+The performance of this combination is shown in the third bar.
+As said this is very safe, therefore the travelled distance is lower than only applying edge detection.
+
+An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
+A region is considered road if there are no edges.
+When there are edges, road classification is used to decide whether it's road.
+This strategy gave us the best results and are shown as the fourth bar.
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\textwidth]{fig/distance_result.png}
+	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
+\end{figure}
+
+This last combination based on edge detection is able to cope with zebra crossings.
+This is shown in figure \ref{zebra}.
+In the first one the road markings are small because the car is far enough.
+Edge detection is therefore able to filter out these markings.
+In the next frame the markings are closer to the car and are too large to be filtered out.
+Because the road classification suggests the markings are road, they are considered road.
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
+  \caption{Zebra crossing filtered out by edge detection.}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
+  \caption{Zebra crossing classified as road.}
+\end{subfigure}
+\caption{Zebra crossing\label{zebra}}
+\end{figure}
 
@@
 This is repeated until we become a stable condition without outliers.
-
 Next, the edges are checked if they aren't on the road.

### 2016-06-15 Administratie/report/paper.tex
update paper

--- 2016-06-15 Administratie/report/paper.tex
@@
 \subsection{Zebra crossings}
+TODO
 
-This last combination based on edge detection is able to cope with zebra crossings.
-This is shown in figure \ref{zebra}.
-In the first one the road markings are small because the car is far enough.
-Edge detection is therefore able to filter out these markings.
-In the next frame the markings are closer to the care and are too large to be filtered out.
-Because the road classification suggests the markings are road, they are considered road.
+\subsection{Combination of classification and edge detection}\label{method-combination}
+
+For each frame the given masks show the projected trajectory of the car.
+Intersections get calculated in order to find the road where it's safe to drive.
+The edges of the edge detection are visualized as blue lines on figure \ref{visualization}.
+The intersections are shown in green.
+The closest intersection to the car is shown in red.
+At this point the maximum speed is derived from the intensity of this pixel.
+
+In order to calculate the edges, both road classification and edge detection are combined. Therefor two strategies were tested:
+\begin{enumerate}
+\item  A region is considered to be road only if it's identified as road by the road classification and there are no edges.
+Both classification and edge detection should indicate a pixel as road in order to be used for safe driving.
+\item A region is considered to be road if there are no edges. When edges are detected, road classification is used to decide what part of the image is road. 
+\end{enumerate}
+
+Both strategies start by removing outliers from the road classification.
+The goal of this is to eliminate the effect an outlier would have when it's close to the car.
+A non-road block is considered an outlier when it has less than 4 non-road block neighbours.
+This is repeated until all outliers are removed.
+
+For the first strategy the edges of the classification are computed.
+These get combined with the edges from the edge detection.
+The closest intersection of any edge with the given mask of the image is used to calculate the maximum speed of the vehicle.
+
+In the second strategy, the edges are checked if they aren't on the road.
+This is done by checking that an edge is on a non-road block of the road classification.
+For safety reasons there is an additional check.
+The above, left and right blocks of an edge should also be considered as non-road.
+When the edge is not classified as road, it is added to a collection.
+The intersections of the edges in the collection and the masks are then computed.
 
@@
   \centering
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
-  \caption{Zebra crossing filtered out by edge detection.}
-\end{subfigure}%
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
-  \caption{Zebra crossing classified as road.}
-\end{subfigure}
-\caption{Zebra crossing\label{zebra}}
+  \includegraphics[width=.9\textwidth]{fig/frame00025.png}
+  \caption{Visualization of a processed frame showing both classification and edge detection}
+  \label{visualization}
 \end{figure}
-
-\subsection{Combination of classification and edge detection}
-
-The combination of road classification and edge detection starts with removing outliers from the road classification.
-The goal of this is to eliminate the effect an outlier would have when it's close to the car.
-A non-road block is considered an outlier when it has less than 4 non-road block neighbours.
-This is repeated until we become a stable condition without outliers.
-
-Next, the edges are checked if they aren't on the road.
-This is done by checking that an edge is on a non-road block of the road classification.
-For safety reasons there is an additional check.
-The above, left and right blocks of an edge should also be considered as non-road.
-
-When the edge is not classified on the road, it is added to a collection.
-The intersections of the edges in the collection and the masks are computed.
-For each frame the masks give the projected trajectory of the car.
-The edges are visualised as blue lines on figure \ref{zebra}.
-The intersections are shown green.
-The closest intersection to the car is shown in red.
-At this point the maximum speed is derived from the intensity.
 
@@
 These scores are used to compare the classification results (higher F-scores indicate better classification).
+
+todo: geen schatten van edge detection meer
 
@@
 \subsection{Driving results}
-To combine road classification and edge detection we should first look at the performance of these two methods separately.
-Therefore, the evaluation was run on each of these two methods.
-The result is shown in figure \ref{distance_result}.
+Driving results get evaluated using a script that calculates the distance a vehicle can drive in one minute, based on the given speed by our system.
+This script includes a ''ground truth'' for each frame in each dataset.
+If our speeds exceed this ground truth, the vehicle crashes and is not able to continue driving.
+
+In section \ref{method-combination} two strategies were presented.
+These use a combination of classification and edge detection.
+First, the evaluation was run separately on these two methods.
+The results are shown in figure \ref{distance_result}.
 The first bar of the histogram shows for every dataset the performance of road classification.
-Because of a crash in the first dataset there can only be travelled 317m.
-The second bar shows for each dataset the distance travelled only relying on edge detection.
-There are no crashes in the four datasets so the results are better.
+Because of a crash in the first dataset there can only be traveled 317m (out of the 1350m).
+The second bar shows for each dataset the distance traveled only relying on edge detection.
+There are no crashes in the four datasets so we can say these results are way better.
+This makes us believe the second combination strategy will be the better one, as it relies.
 
-A first and very safe way to combine these two methods is the following.
-A region is considered road only if it's identified as road by the road classification and there is no edge.
-The performance of this combination is shown in the third bar.
-As said this is very safe, therefore the travelled distance is lower than only applying edge detection.
+TODO: maximum met edge detection voor elke dataset? 1350,
 
-An other way to make a combination, witch still results in zero crashes is to focus on edge detection.
+The first way of combining the two methods is very safe and thus results in no crashes over the four datasets.
+A region is considered to be road only if it's identified as road by the road classification and if there are no edges.
+Both classification and edge detection should indicate a pixel as road in order to be used for safe driving.
+The performance of this combination is shown in the third bar of figure \ref{distance_result}.
+As said, this strategy is very safe, giving up a lot of traveled distance in favor of safety.
+
+The second strategy for combination also results in zero crashes over the four datasets.
+This strategy tries to focus on edge detection.
 A region is considered road if there are no edges.
-When there are edges, road classification is used to decide whether it's road.
-This strategy gave us the best results and are shown as the fourth bar.
+When there are edges, road classification is used to decide whether it's road or not.
+This strategy gave us the best results, as shown in figure \ref{distance_result} as the fourth bar.
 \begin{figure}[ht]

### 2016-06-15 Administratie/report/paper.tex

@@
 
-This last combination based on edge detection is able to cope with zebra crossings.
-This is shown in figure \ref{zebra}.
-In the first one the road markings are small because the car is far enough.
-Edge detection is therefore able to filter out these markings.
-In the next frame the markings are closer to the care and are too large to be filtered out.
-Because the road classification suggests the markings are road, they are considered road.
+As an extra, this strategy is also able to cope better with zebra crossings than the first strategy. This is shown in figure \ref{zebra}. In the first figure (\ref{zebra-1}) TODO
 
@@
   \caption{Zebra crossing filtered out by edge detection.}
+  \label{zebra-1}
 \end{subfigure}%
@@
   \caption{Zebra crossing classified as road.}
+  \label{zebra-2}
 \end{subfigure}

### 2016-06-15 Headers/io.hpp
bugfix trainingsdata useLBP and useColor mix-up

@@ -34,10 +34,10 @@ namespace io {
     const int innerMargin = 64;
     const int blkSize = 16; // 8,16,32
     const int lbpRadius = 1;
-    const bool includeMarks = true;
+    const bool includeMarks = false;
     const bool trainAuto = false;
     const bool useLBP = true; // Make sure to use either color, LBP or both
-    const bool useColor = false;
+    const bool useColor = true;
 
     /** SVM parameters for when not training automatically **/
     const double C = 0.1;
@@ -216,7 +216,7 @@ namespace io {
 	    string fnLbl = dirOutput+frameName+labelsPostfix;
 	    string fnFv = dirOutput+frameName+(useLBP?lbpStr:"")+(useColor?colorStr:"")+featVecsPostfix;
 	    if(file_exists(fnLbl) && file_exists(fnFv)) {
-			cout<<"Reading trainingsdata from output of frame "<<frameName<<endl;
+			cout<<"Reading trainingsdata from "<<fnFv<<endl;
 		 	ifstream ifsLbl(fnLbl.c_str());
 		 	ifstream ifsFv(fnFv.c_str());
 

### 2016-06-15 Headers/lbpfeaturevector.hpp

@@ -176,6 +176,8 @@ public:
 	* along with a copy of the frame with the blocks drawn on top of it.
 	*/
 	void processFrame(string fnFrame, Mat& img, Mat& featVectors, bool useColor=io::useColor, bool useLBP=io::useLBP, bool isTrainingsdata=false){
+	// void processFrame(string fnFrame, Mat& img, Mat& featVectors, bool useColor=true, bool useLBP=true, bool isTrainingsdata=false){
+
     	cout<<"Start processFrame "<<fnFrame+(useLBP?io::lbpStr:"")+(useColor?io::colorStr:"")<<endl;
     	if(!useColor && !useLBP){
     		cerr<<"You are using neither color or LBP for featurvectors, aka nothing!"<<endl;
@@ -193,13 +195,6 @@ public:
 
 		int imWidth=img.cols;
 		int imHeight=img.rows;
-
-		// io::checkDir(io::dirOutput);
-		// size_t pos1 = fnFrame.find_last_of("/")+1;
-		// if(pos1 == string::npos) pos1=0;
-		// size_t pos2 = fnFrame.find_last_of(".");
-		// if(pos2 == string::npos) pos2=fnFrame.size();
-		// string frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
 
 
 		// Number of blocks in width and height

### 2016-06-15 maxSpeed.cpp

@@ -66,10 +66,11 @@ bool isRoad(vector<int> & roadRegions, int blksInWidth, int frameRows, int frame
     }
 }
 
-void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames, vector<vector<int> > roadRegions, vector<double> speeds, string dirOutputFrames) {
+void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames, vector<vector<int> > roadRegions, vector<double> speeds, string dirOutputFrames, string& results) {
     int crash = 0;
 
     int thresh = 255;
+    stringstream ssOut;
     for(int i = 0; i<masks.size(); i++) {
         Mat col = cv::Mat::ones(roads[i].rows, 1, roads[i].type());
         Mat rows = cv::Mat::ones(masks[i].rows-roads[i].rows, masks[i].cols, roads[i].type());
@@ -252,20 +253,21 @@ void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames
         //print result
         //cout << "FRAME " << i << ": " << laagsteSnelheid << " (" << speeds[i] << ") -> " << laagsteSnelheid-speeds[i] << endl;
 
-        stringstream ss2;
-        ss2 << setfill('0') << std::setw(5) << i << " " << laagsteSnelheid;
-        cout << ss2.str() << endl;
+        ssOut << setfill('0') << std::setw(5) << i << " " << laagsteSnelheid <<endl;
 
         /// Show in a window
         /*namedWindow( "Max Speed", CV_WINDOW_AUTOSIZE );
         imshow( "Max Speed", dst );
         waitKey(0);*/
         io::checkDir(dirOutputFrames);
-        stringstream ssOut;
-        ssOut << dirOutputFrames << "/frame" << setfill('0') << std::setw(5) << i << ".png";
-        imwrite(ssOut.str(),dst);
+        stringstream ssFilename;
+        ssFilename << dirOutputFrames << "/frame" << setfill('0') << std::setw(5) << i << ".png";
+        imwrite(ssFilename.str(),dst);
     }
-    cout << "CRASHES: " << crash << endl;
+    ssOut << "CRASHES: " << crash << endl;
+
+    results = ssOut.str();
+    cout<<results<<endl;
 }
 
 vector<Mat> detectLines(vector<Mat> & masks, vector<Mat> & frames){
[…]

### 2016-06-15 trainingsdata.cpp

@@ -232,7 +232,7 @@ void parameterIterationTraining(bool relabel=true){
 		    Mat featureVectors;
 
 		    io::buildFrameName(buffer,frameName,dataset,fIdx,innerMargin,blkSize,includeMarks);
-		    fv.processFrame(frameName,frame,featureVectors,useLBP,useColor,true);
+		    fv.processFrame(frameName,frame,featureVectors,useColor,useLBP,true);
 
 		    vector<bool> isRoad;
 		    generateLabels(frameName,isRoad,featureVectors.rows);
@@ -339,35 +339,8 @@ void parameterIterationTraining(bool relabel=true){
 	csv.close();
 }
 
-void fullTraining(bool relabel=false){
- //    LbpFeatureVector fv(io::outerMargin,io::innerMargin);
-
-	// vector<Mat> frames;
-	// string frameRegex = "frame%05d.png";
-	// for(int s=0; s<sizeof(io::datasetFolders)/sizeof(string); s++) {
-	// 	io::readImages(io::datasetFolders[s],frameRegex,frames);
-	// 	for(int i=0;i<frames.size();i+=5){
-	// 		char buffer[30];
-	// 		sprintf(buffer,("%02d"+frameRegex).c_str(),s+1,i);
-	// 		string fnFrame = string(buffer);
-	// 		Mat featureVectors;
-	// 		fv.processFrame(fnFrame,frames[i],featureVectors,true,true,true);
-
-	// 		string frameName;
-	// 		fileToFrameName(fnFrame,frameName);
-
-	// 		vector<bool> isRoad;
-	// 		generateLabels(frameName,isRoad,featureVectors.rows);
-	// 		if(relabel)	guiLabeling(fv,frames[i],isRoad);
-
-	// 		saveFrameOutput(frameName,frames[i],featureVectors,isRoad);
-	// 	}
-	// }
-}
-
 int main (int argc, char** argv){
 	parameterIterationTraining();
-	// fullTraining();
 
     return 0;
 }

### 2016-06-15 trainingsdata.cpp
new trainingsdata

@@ -118,7 +118,7 @@ void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
 	}
 }
 
-void saveFrameOutput(const string frameName, const Mat& frame, const Mat& featureVectors, vector<bool>& isRoad, bool useLBP, bool useColor){
+void saveFrameOutput(const string frameName, const Mat& frame, const Mat& featureVectors, vector<bool>& isRoad, bool useColor, bool useLBP){
 	vector<short> labels(isRoad.size());
 	for(int i=0;i<isRoad.size();i++)
 		labels[i] = (isRoad[i]? 1 : -1);
@@ -199,7 +199,7 @@ void parameterIterationTraining(bool relabel=true){
     int frameInterval = 10;
     // int frameStopIdx = io::FRAME_MAX_IDX;
     int frameStopIdx = 50;
-	bool trainAuto = false; // Whether or not to use automatic training for SVM
+	bool trainAuto = true; // Whether or not to use automatic training for SVM
 	bool includeMarksVals[] = {false,true};
 	bool useColorVals[] = {false,true,true};
 	bool useLBPVals[] = {true,false,true};
@@ -232,7 +232,7 @@ void parameterIterationTraining(bool relabel=true){
 		    Mat featureVectors;
 
 		    io::buildFrameName(buffer,frameName,dataset,fIdx,innerMargin,blkSize,includeMarks);
-		    fv.processFrame(frameName,frame,featureVectors,useColor,useLBP,true);
+		    fv.processFrame(frameName,frame,featureVectors,useColor,useLBP);
 
 		    vector<bool> isRoad;
 		    generateLabels(frameName,isRoad,featureVectors.rows);
@@ -243,7 +243,7 @@ void parameterIterationTraining(bool relabel=true){
 			    guiLabeling(fv,frameWithBlocks,isRoad);
 			}
 
-		    saveFrameOutput(frameName,frameWithBlocks,featureVectors,isRoad,useLBP,useColor);
+		    saveFrameOutput(frameName,frameWithBlocks,featureVectors,isRoad,useColor,useLBP);
 
 		    destroyAllWindows();
 		}

### 2016-06-15 Administratie/report/paper.tex
update paper

@@
 \subsection{Precision, recall, F-scores}
-
 TODO: Bjorn
@@
 In order to determine the maximum speed of a vehicle two image processing techniques are used: SVM classification and edge detection.
-We first try to rely on edge detection to find the contours of the lanes where driving is permitted (i.e.
-where no obstacles are found).
-We estimate the reliability of the edge detection and incorporate road detection using the SVM classifier when needed.
+In order to combine the best of both techniques, two strategies are discussed in section \ref{method-combination}.
 
 For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either train or test the solution.
-To train out SVM classifier, we used a function of the OpenCV SVM library which selects the optimal parameters using cross-validation.
 
@@
 We compare classification results varying these parameters for the feature vector in section \ref{results-svm-classification}.
+To train our SVM classifier, we use a function of the OpenCV SVM library which selects the optimal parameters using cross-validation.
 
@@
 Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection \cite{canny}.
-However, there are a lot of edges on the road itself, thinking of road markings and cobbles.
+However, there might be a lot of edges on the road itself (think of road markings and cobbles).
 The edges on the road, considered as noise, and the edges of the road are not unambiguously separable.
@@
 
-Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{sreedhar2012enhancement}.
+Eroding \cite{sreedhar2012enhancement} is a morphology operator to make the objects on the foreground, which are the brightest, smaller.
 Figure \ref{line_zoom_eroded} shows an example of a thinned white line.
@@
 This will falsify the eventual edge detection.
-In this example, the observed car is reported to close.
+In this example, the observed car is reported too close.
 In order to restore the original measures, dilation is executed.
@@
 When the algorithm is configured too sensitive, some imperfections on the road will be detected as edges.
-On the other hand, a certain level of sensitivity is necessary to detect less clear road edges.
+On the other hand, a certain level of sensitivity is necessary to detect road edges that are  less clear.
 For example, a transition from road to dirt has a very small distinction of color and brightness, as seen in \ref{transition_road_dirt_eroded}.
@@
 
-\subsection{Zebra crossings}
-TODO
-
 \subsection{Combination of classification and edge detection}\label{method-combination}
 
-For each frame the given masks show the projected trajectory of the car.
-Intersections get calculated in order to find the road where it's safe to drive.
+In order to calculate the maximum speed a vehicle can have in a particular frame, we use a combination of road classification and edge detection.
+Like that, we can detect where it is safe to drive.
+For each frame in the dataset, some masks are given that show the projected trajectory of the car in that particular frame.
+Intersections between this mask and the ''safe zone'' in the frame get calculated in order to find the maximum safe speed for each frame.
 The edges of the edge detection are visualized as blue lines on figure \ref{visualization}.
@@
 
-In order to calculate the edges, both road classification and edge detection are combined. Therefor two strategies were tested:
+In order to calculate the edges, two different strategies were tested:
 \begin{enumerate}
@@
 Both strategies start by removing outliers from the road classification.
-The goal of this is to eliminate the effect an outlier would have when it's close to the car.
-A non-road block is considered an outlier when it has less than 4 non-road block neighbours.
+The goal of this is to eliminate the effect that outliers would have when it's close to the car.
+A non-road block is considered an outlier when it has less than 4 non-road neighbors.
 This is repeated until all outliers are removed.
 
-For the first strategy the edges of the classification are computed.
+For the first strategy, the edges of the area classified as road are computed.
 These get combined with the edges from the edge detection.
@@
 
-In the second strategy, the edges are checked if they aren't on the road.
-This is done by checking that an edge is on a non-road block of the road classification.
+In the second strategy, the edges are first calculated using edge detection.
+Each edge is then checked whether it lies on the road or not using road classification.
 For safety reasons there is an additional check.
 The above, left and right blocks of an edge should also be considered as non-road.
-When the edge is not classified as road, it is added to a collection.
+When an edge lies on a part of the frame that is not classified as road, it is added to a collection.
 The intersections of the edges in the collection and the masks are then computed.
@@
 These scores are used to compare the classification results (higher F-scores indicate better classification).
-
-todo: geen schatten van edge detection meer
 
@@
 
-TODO: maximum met edge detection voor elke dataset? 1350,
-
 The first way of combining the two methods is very safe and thus results in no crashes over the four datasets.
@@
 	\centering
-	\includegraphics[width=\textwidth]{fig/distance_result.png}
+	\includegraphics[width=\textwidth]{fig/distance_result}
 	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}

### 2016-06-15 maxSpeed.cpp
Safety zone of two blocks

@@ -46,11 +46,21 @@ bool isRoad(vector<int> & roadRegions, int blksInWidth, int frameRows, int frame
 
     // most left and right column are never considered as road, in order to limit the speed when
     // reaching the edge of the frame (i.e. necessary on the roundabout in Dataset 01)
-    if(x < io::blkSize || x >= frameCols - io::blkSize) return false;
+    if(x < io::blkSize * 2 || x >= frameCols - (io::blkSize * 2)) return false;
 
     // the lowest row is always considered as road
     if(y >= frameRows - io::blkSize) return true;
 
+    for(int k = -2; k <= 2; k++) {
+        if(
+            roadRegions[j-(blksInWidth * 2)+k] != 1 ||
+            roadRegions[j-blksInWidth+k] != 1 ||
+            roadRegions[j+k] != 1
+        ) return false;
+    }
+
+    return true;
+/*
     if(
         roadRegions[j] == 1 &&
         // above
@@ -58,12 +68,10 @@ bool isRoad(vector<int> & roadRegions, int blksInWidth, int frameRows, int frame
         roadRegions[j-blksInWidth] == 1 &&
         roadRegions[j-blksInWidth+1] == 1 &&
 
-        // left & right
-        roadRegions[j-1] == 1 &&
-        roadRegions[j+1] == 1
+
     ) {
         return true;
-    }
+    }*/
 }
 
 void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames, vector<vector<int> > roadRegions, vector<double> speeds, string dirOutputFrames, string& results) {
@@ -341,8 +349,16 @@ int main(int argc, char** argv){
         iss >> speed;
         speeds.push_back(speed);
     }
-    vector<Mat> masks = read_images(dirDataset,"mask%05d.png");
-    vector<Mat> frames = read_images(dirDataset,"frame%05d.png");
+    vector<Mat> masks_all = read_images(dirDataset,"mask%05d.png");
+    vector<Mat> frames_all = read_images(dirDataset,"frame%05d.png");
+
+    vector<Mat> masks;
+    vector<Mat> frames;
+    for(int i=0; i < 30; i++) {
+        masks.push_back(masks_all[i]);
+        frames.push_back(frames_all[i]);
+    }
+
 
     vector<Mat> roads = detectLines(masks,frames);
     cout << "lines done"<< endl;

### 2016-06-15 Administratie/report/paper.tex
update report

@@
 
-%
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
-%
-
 \maketitle
@@
 This is the abstract: TODO
-\keywords{SVM classification, Local Binary Patterns, line detection}
+\keywords{SVM classification, Local Binary Patterns, edge detection}
 \end{abstract}
@@
 \section{Introduction}
+In recent years, a lot of advanced electronics have been used to help automate driving.
+It is already common for some cars to have (semi-)automated parking \cite{Pohl2006} and adaptive cruise control \cite{Jurgen}.
+Also, a lot of safety features are being developed to increase environmental awareness of cars: pedestrian detection \cite{Gavrila2000}, automated detection of speed limit signs \cite{Torresen2004}, etc.
+
+A feature that could be realized using the electronics that are already available in some cars, is automated maximum speed detection.
+This speed detection would simulate the braking distance towards the nearest obstacle and inform the user.
+There are various ways to achieve this.
+Commonly used systems in adaptive cruise control are lidar and radar.
+The choice of sensor presents classic design tradeoffs \cite{Jones2001}.
+Lidar  is  less  expensive  to  produce  and  easier  to package  but  performs  poorly  in  rain  and  snow.
+Another problem is that accumulations of mud, dust or snow on the car
+can block lidar beams. Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver’s ability to see down to 10  meters  or  less.
+
+Because of the problems with both this systems, we propose to use a detection system based on camera images.
+These images present a few advantages over a system using lidar or radar.
+First of all cameras are way cheaper to install.
+On top of that they see color (whereas lidar only sees gray scale in the infrared spectrum) and are able to gain very high resolution for more distant objects.
+There are several downsides to using cameras.
+They need light to work properly.
+This becomes a problem at night, because cameras can only rely on the headlights of the car.
+Secondly computer vision systems using camera images can use a lot op CPU power.
 TODO
@@
 
-\subsection{Line detection}
-TODO: Mathias
-
 \section{Method}
-In order to determine the maximum speed of a vehicle two image processing techniques are used: SVM classification and edge detection.
-In order to combine the best of both techniques, two strategies are discussed in section \ref{method-combination}.
+In order to determine the maximum speed of a vehicle, two image processing techniques are used: SVM classification and edge detection.
+To combine the best of both techniques, two strategies are discussed in section \ref{method-combination}.
 
@@
 A linear kernel is faster, but the RBF kernel will outperform the linear kernel in case the data is not linearly separable.
-We tested and compared the results using each kernel to decide which is best fit for our classification problem.
+We tested and compared the results using each kernel to decide which is the best fit for our classification problem.
 
@@
 For a training set (a subset of the datasets available) each frame is divided into blocks.
-Each block must be manually labelled with the expected value for classification.
-For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset (i=0,10,...,40).
+Each block must be manually labeled with the expected value for classification.
+For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset ($i=0,10,...,40$).
 Figure \ref{fig:frame-training} shows a visualization of pre-labeled blocks in a frame.
@@
 \subsection{Road marks}
-The training dataset was labelled considering blocks containing road marks once as part of the road class and once as part of the non-road class.
-Initially road marks were labelled as non-road, in order to get a good classification of road (i.e.
+The training dataset was labeled twice, once considering blocks containing road marks as part of the road class and once as part of the non-road class.
+Initially road marks were labeled as non-road, in order to get a good classification of road (i.e.
 different types of asphalt).
-As these road marks aren't really considered an obstacle for a car driver, ideally they should be classified as road.
+But as these road marks aren't really considered an obstacle for a car driver, ideally they should be classified as road.
 This makes up for a more difficult classification problem.
@@
 Like that, we can detect where it is safe to drive.
-For each frame in the dataset, some masks are given that show the projected trajectory of the car in that particular frame.
+For each frame in the dataset, a mask is given that shows the projected trajectory of the car in that frame.
 Intersections between this mask and the ''safe zone'' in the frame get calculated in order to find the maximum safe speed for each frame.
@@
 In the second strategy, the edges are first calculated using edge detection.
-Each edge is then checked whether it lies on the road or not using road classification.
+Each edge is then checked for whether it lies on the road or not using road classification.
 For safety reasons there is an additional check.
@@
 
-Of the four datasets, different combinations of training sets were selected (each time shown as x-axis on the graphs). The trained classifier was then tested on the remaining datasets.
+Different combinations of training sets were selected of the four given datasets (training sets are each time shown on the x-axis on the graphs).
+The trained classifier is then tested on the remaining datasets.
 
@@
 better accuracy).
-This is why prefer the block size of 16x16px.
+This is why we prefer the block size of 16x16px.
 
@@
 
-If we do not include road marks in the training data, we get slightly better results when combining LBP and color for the feature vectors (as shown in figure \ref{fig:16_wo_roadmarks}).
-If on the other hand we do include these road marks, using only LBP shows equal or better results than the combination of LBP and color (as shown in figure \ref{fig:16_wi_roadmarks}).
+If we do not include road marks in the training data, we get slightly better results when combining LBP and color for the feature vectors, as shown in figure \ref{fig:16_wo_roadmarks}.
+If on the other hand we do include these road marks, using only LBP shows equal or better results than the combination of LBP and color, as shown in figure \ref{fig:16_wi_roadmarks}.
 
@@
 
-Considering the results of the different scenarios, we can conclude that training on dataset four results in poorer classification overall.
+Considering the results of the different scenarios, we can conclude that training on the fourth dataset results in poorer classification overall.
 This is because another type of road is used in this dataset, conflicting with the more common road type present in the other datasets.
@@
 In section \ref{method-combination} two strategies were presented.
-These use a combination of classification and edge detection.
+These use a combination of both classification and edge detection.
 First, the evaluation was run separately on these two methods.
@@
 The first bar of the histogram shows for every dataset the performance of road classification.
-Because of a crash in the first dataset there can only be traveled 317m (out of the 1350m).
-The second bar shows for each dataset the distance traveled only relying on edge detection.
-There are no crashes in the four datasets so we can say these results are way better.
-This makes us believe the second combination strategy will be the better one, as it relies.
+Because of a crash in the first dataset there can only be traveled 317 meters (out of the total 1350 meters in this dataset).
+The second bar shows the distance traveled only relying on edge detection for each dataset.
+There are no crashes in the four datasets so we can say these results are way better than the ones only using classification.
+This makes us believe the combination strategy that relies more on edge detection (strategy 2) will be the better one.
 

### 2016-06-15 Administratie/report/paper.tex
update paper

@@
 \section{Introduction}
+The goal of this paper is to determine a maximum speed for a moving vehicle using only camera images and information from the steering column.
+
 In recent years, a lot of advanced electronics have been used to help automate driving.
@@
 
-A feature that could be realized using the electronics that are already available in some cars, is automated maximum speed detection.
+A additional feature that can be realized using the electronics that are already available in some cars, is automated maximum speed detection.
 This speed detection would simulate the braking distance towards the nearest obstacle and inform the user.
@@
 Commonly used systems in adaptive cruise control are lidar and radar.
-The choice of sensor presents classic design tradeoffs \cite{Jones2001}.
+The choice of sensor presents classic design trade-offs \cite{Jones2001}.
 Lidar  is  less  expensive  to  produce  and  easier  to package  but  performs  poorly  in  rain  and  snow.
-Another problem is that accumulations of mud, dust or snow on the car
-can block lidar beams. Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver’s ability to see down to 10  meters  or  less.
+Another problem is that accumulations of mud, dust or snow on the car can block lidar beams.
+Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver’s ability to see down to 10  meters  or  less.
 
-Because of the problems with both this systems, we propose to use a detection system based on camera images.
+Because of the problems with both these systems, we propose to use a detection system based on camera images.
 These images present a few advantages over a system using lidar or radar.
 First of all cameras are way cheaper to install.
-On top of that they see color (whereas lidar only sees gray scale in the infrared spectrum) and are able to gain very high resolution for more distant objects.
+On top of that they see color and are able to gain very high resolution for more distant objects.
 There are several downsides to using cameras.
-They need light to work properly.
-This becomes a problem at night, because cameras can only rely on the headlights of the car.
-Secondly computer vision systems using camera images can use a lot op CPU power.
-TODO
+They require light to work properly, which becomes a problem for example at night.
+Secondly computer vision systems using camera images can use a lot of CPU power.
+Nonetheless cameras provide a cheap and accessible platform for immediate applications in real-world scenarios.
+
+Section 2 discusses some relevant background information mostly concerning Local Binary Patterns and SVM classification.
+Section 3 describes the methods used in order to achieve the best results.
+The results are then discussed in section 4.
+In the last section an overview of the most important results is presented.
 

### 2016-06-15 Administratie/report/paper.tex
update report

@@
 \documentclass[runningheads,a4paper]{llncs}
-
+\usepackage[english]{babel}
 \usepackage{amssymb}
@@
 \captionsetup{compatibility=false}
+\usepackage{booktabs}
+\usepackage{float}
 
@@
 \newcommand{\keywords}[1]{\par\addvspace\baselineskip
+\newcommand\fnumber{\mathop{\mbox{$F$-$score$}}}
 \noindent\keywordname\enspace\ignorespaces#1}
@@
 \begin{abstract}
-This is the abstract: TODO
+This is the abstract: The goal of this paper it to determine the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column.
+We construct a system that combines SVM classification and edge detection to foresee possible obstacles on the given trajectory of the car.
+The maximum speed is computed so that braking in time will avoid collision with those obstacles.
+TODO: hoe goed is dit systeem?
 \keywords{SVM classification, Local Binary Patterns, edge detection}
@@
 \section{Introduction}
-The goal of this paper is to determine a maximum speed for a moving vehicle using only camera images and information from the steering column.
+This paper determines the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column.
+We construct a system that combines SVM classification and edge detection to foresee possible obstacles on the given trajectory.
+The maximum speed is computed so that braking in time will avoid collision with those obstacles.
 
@@
 
-A additional feature that can be realized using the electronics that are already available in some cars, is automated maximum speed detection.
+An additional feature that can be realized using the electronics that are already available in these cars, is automated maximum speed detection.
 This speed detection would simulate the braking distance towards the nearest obstacle and inform the user.
 There are various ways to achieve this.
-Commonly used systems in adaptive cruise control are lidar and radar.
+Commonly used sensors in adaptive cruise control are lidar and radar.
 The choice of sensor presents classic design trade-offs \cite{Jones2001}.
@@
 Another problem is that accumulations of mud, dust or snow on the car can block lidar beams.
-Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver’s ability to see down to 10  meters  or  less.
+Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver's ability to see down to 10 meters or less.
 
-Because of the problems with both these systems, we propose to use a detection system based on camera images.
-These images present a few advantages over a system using lidar or radar.
+Because of the problems with both these systems, this paper proposes the use of a detection system based on camera images.
+These images present some advantages over a system using lidar or radar.
 First of all cameras are way cheaper to install.
@@
 Nonetheless cameras provide a cheap and accessible platform for immediate applications in real-world scenarios.
+On top of that, cameras could be used in combination with lidar and radar to provide even better results.
 
+\subsection{Overview}
 Section 2 discusses some relevant background information mostly concerning Local Binary Patterns and SVM classification.
@@
 The results are then discussed in section 4.
-In the last section an overview of the most important results is presented.
+In the last section, an overview of the most important results is presented.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centering
@@
 
-\subsection{Precision, recall, F-scores}
-TODO: Bjorn
-
 \section{Method}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \centering

### 2016-06-15 Administratie/report/paper.tex
Paper classifier results

--- 2016-06-15 Administratie/report/paper.tex
@@
 For each combination of parameters we train and test on different combinations of datasets.
-For each combination we calculate the F-scores.
+For each combination we calculate the F1-scores \cite.
 These scores are used to compare the classification results (higher F-scores indicate better classification).
 
-Different combinations of training sets were selected of the four given datasets (training sets are each time shown on the x-axis on the graphs).
+Different combinations of training sets were selected of the four given datasets (training sets used are each time shown on the x-axis on the graphs).
 The trained classifier is then tested on the remaining datasets.
@@
 As previously mentioned in section \ref{methods-kernels}, there are two SVM kernels we can use for this classification problem.
-On average, we tend to get a nearly 11\% increase in F-scores using RBF kernel instead of linear kernel.
+On average, we tend to get a nearly 10\% increase in F-scores using an RBF kernel instead of a linear kernel.
 It is safe to say the RBF kernel outperforms the linear kernel for this classification problem in all of our test scenarios.
@@
 In general, block sizes 16x16px and 32x32px tend to show the better results.
-As an example we present the F-scores for LBP and color with road marks in figure \ref{fig:lbp-col-wi-rm}.
-Other combinations follow the same trend.
+As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}.
+Other combinations follow the same trend, although 32x32px sometimes performs slightly better.
 For the goal of this application, a smaller block size will be more practical (i.e.
@@
 \centering
-\includegraphics[width=\textwidth]{fig/RBF_LBP_Col_wi_RM.png}
-\caption{F-scores for LBP and color with road marks using RBF kernel}
-\label{fig:lbp-col-wi-rm}
+\includegraphics[width=\textwidth]{fig/RBF_Col_wi_RM.png}
+\caption{F-scores for color with road marks using RBF kernel}
+\label{fig:col-wi-rm}
 \end{figure}

### 2016-06-15 Headers/io.hpp

@@ -36,7 +36,7 @@ namespace io {
     const int lbpRadius = 1;
     const bool includeMarks = true;
     const bool trainAuto = true;
-    const bool useLBP = true; // Make sure to use either color, LBP or both
+    const bool useLBP = false; // Make sure to use either color, LBP or both
     const bool useColor = true;
 
     /** SVM parameters for when not training automatically **/

### 2016-06-15 trainingsdataFTT.cpp

@@ -1,0 +1,346 @@
+#include <string>
+#include <stdio.h>
+// #include <time.h>
+#include <set>
+#include "Headers/io.hpp"
+#include "Headers/lbpfeaturevector.hpp"
+#include "Headers/svm.hpp"
+
+using namespace std;
+using namespace cv;
+
+void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
+	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
+	if(io::file_exists(fnLabels)){
+		io::readFrameLabels(fnLabels,isRoad);
+	}else{
+		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
+		isRoad = vector<bool>(totalBlocks);
+
+		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
+			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
+	}
+}
+
+static void onMouse( int event, int x, int y, int, void* param){
+    if( event != EVENT_LBUTTONDOWN )
+        return;
+
+    int* coordinates = (int*) param;
+    coordinates[0] = x;
+    coordinates[1] = y;
+}
+
+void fileToFrameName(string fnFrame, string& frameName){
+	size_t pos1 = fnFrame.find_last_of("/")+1;
+	if(pos1 == string::npos) pos1=0;
+	size_t pos2 = fnFrame.find_last_of(".");
+	if(pos2 == string::npos) pos2=fnFrame.size();
+	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
+}
+
+void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
+	/* Initial configuration */
+    ostringstream strBldr; // Stringbuilder
+	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
+	imWidth = img.cols;
+	imHeight = img.rows;
+	outerMargin = fv.getOuterMargin();
+	innerMargin = fv.getInnerMargin();
+	blkX = blkY = outerMargin;
+	blkSize = fv.getBlkSize();
+
+	/* Draw initial block labeling */
+	for(int blkIdx=0; 
[…]

### 2016-06-15 trainingsdataTFT.cpp

@@ -1,0 +1,346 @@
+#include <string>
+#include <stdio.h>
+// #include <time.h>
+#include <set>
+#include "Headers/io.hpp"
+#include "Headers/lbpfeaturevector.hpp"
+#include "Headers/svm.hpp"
+
+using namespace std;
+using namespace cv;
+
+void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
+	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
+	if(io::file_exists(fnLabels)){
+		io::readFrameLabels(fnLabels,isRoad);
+	}else{
+		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
+		isRoad = vector<bool>(totalBlocks);
+
+		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
+			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
+	}
+}
+
+static void onMouse( int event, int x, int y, int, void* param){
+    if( event != EVENT_LBUTTONDOWN )
+        return;
+
+    int* coordinates = (int*) param;
+    coordinates[0] = x;
+    coordinates[1] = y;
+}
+
+void fileToFrameName(string fnFrame, string& frameName){
+	size_t pos1 = fnFrame.find_last_of("/")+1;
+	if(pos1 == string::npos) pos1=0;
+	size_t pos2 = fnFrame.find_last_of(".");
+	if(pos2 == string::npos) pos2=fnFrame.size();
+	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
+}
+
+void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
+	/* Initial configuration */
+    ostringstream strBldr; // Stringbuilder
+	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
+	imWidth = img.cols;
+	imHeight = img.rows;
+	outerMargin = fv.getOuterMargin();
+	innerMargin = fv.getInnerMargin();
+	blkX = blkY = outerMargin;
+	blkSize = fv.getBlkSize();
+
+	/* Draw initial block labeling */
+	for(int blkIdx=0; 
[…]

### 2016-06-15 trainingsdataTTT.cpp

@@ -1,0 +1,346 @@
+#include <string>
+#include <stdio.h>
+// #include <time.h>
+#include <set>
+#include "Headers/io.hpp"
+#include "Headers/lbpfeaturevector.hpp"
+#include "Headers/svm.hpp"
+
+using namespace std;
+using namespace cv;
+
+void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
+	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
+	if(io::file_exists(fnLabels)){
+		io::readFrameLabels(fnLabels,isRoad);
+	}else{
+		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
+		isRoad = vector<bool>(totalBlocks);
+
+		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
+			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
+	}
+}
+
+static void onMouse( int event, int x, int y, int, void* param){
+    if( event != EVENT_LBUTTONDOWN )
+        return;
+
+    int* coordinates = (int*) param;
+    coordinates[0] = x;
+    coordinates[1] = y;
+}
+
+void fileToFrameName(string fnFrame, string& frameName){
+	size_t pos1 = fnFrame.find_last_of("/")+1;
+	if(pos1 == string::npos) pos1=0;
+	size_t pos2 = fnFrame.find_last_of(".");
+	if(pos2 == string::npos) pos2=fnFrame.size();
+	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
+}
+
+void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
+	/* Initial configuration */
+    ostringstream strBldr; // Stringbuilder
+	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
+	imWidth = img.cols;
+	imHeight = img.rows;
+	outerMargin = fv.getOuterMargin();
+	innerMargin = fv.getInnerMargin();
+	blkX = blkY = outerMargin;
+	blkSize = fv.getBlkSize();
+
+	/* Draw initial block labeling */
+	for(int blkIdx=0; 
[…]

### 2016-06-15 Administratie/report/paper.tex
update report
kleine aanpassingen
Afwerking tekst

@@
 For each combination of parameters we train and test on different combinations of datasets.
-For each combination we calculate the F1-scores \cite.
+For each combination we calculate the F1-scores \cite{Powers2007}.
 These scores are used to compare the classification results (higher F-scores indicate better classification).
@@
 Other combinations follow the same trend, although 32x32px sometimes performs slightly better.
-For the goal of this application, a smaller block size will be more practical (i.e.
-better accuracy).
-This is why we prefer the block size of 16x16px.
+This is why we conclude the block size of 32x32px to be a more reliable choice.
 
@@
 
-\subsubsection{LBP}
-The use of LBP is crucial to our implementation in order to get a viable classifier for road detection.
-If we compare results (block size 16x16px) of feature vectors on LBP only to the combination of LBP and color, we get different results depending on whether or not we include road marks.
+\subsubsection{Feature vector values}
+The use of LBP represents no significant gains in order to get a viable classifier for road detection.
+If we compare results (block size 32x32px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
+Therefore, in the case of our classification problem, we do not add LBP to the feature vector.
 
-If we do not include road marks in the training data, we get slightly better results when combining LBP and color for the feature vectors, as shown in figure \ref{fig:16_wo_roadmarks}.
-If on the other hand we do include these road marks, using only LBP shows equal or better results than the combination of LBP and color, as shown in figure \ref{fig:16_wi_roadmarks}.
+The best results are found when using color histogram values for the feature vector.
+This implies the combinations of color values are sufficient to distinguish the road from other surroundings.
+A possible explanation for this phenomenon is that when asphalt and road marks are considered to be road, color is a much more consistent factor to weigh in to the classification.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{fig/16_wo_roadmarks.png}
-\caption{Comparision of F-scores for 16x16px block size without road marks using RBF kernel }
-\label{fig:16_wo_roadmarks}
-\end{figure}
+\textbf{TODO} grafiek wo/wi RM
 
-A possible explanation for this phenomenon is that when only asphalt is considered to be road, color is a much more consistent factor to weigh in to the classification.
-However when including road marks as part of the road, color intensity is no longer an added value and even tends to mislead the classifier.
+\subsubsection{Road marks}
+Including road marks as part of the road gets better results for the classification problem.
+For our application this is more practical, because road marks should not be considered to be obstacles.
 
-Overall we can conclude that using only LBP with road marks shows the optimal results for our classification problem.
+Overall we can conclude that using only color with road marks shows the optimal results for our classification problem.
 
+TODO figure aanpassen
 \begin{figure}[h]
@@
 \item RBF kernel
-\item Block size 16x16px
-\item Using only LBP histogram values for the feature vectors
+\item Block size 32x32px
+\item Using only color histogram values for the feature vectors
 \item Including road marks as part of the road  
@@
 
-\begin{figure}[t]
+\begin{figure}[H]
 \centering
@@
 \section{Conclusion}
+In this paper we combined SVM classification with edge detection in order to calculate the maximum speed a vehicle can drive, considering possible obstacles on the given trajectory.
+
+The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 32x32px in an image.
+These results tend to get better when we include road marks as part of the road in the training data.
+This is also optimal for use in our application as we don't want to detect road marks as obstacles.
+
+To perform edge detection, the frame is first treated with morphological processing operations to remove insignificant details.
+In our application we used Canny Edge Detection to detect the side of the road and possible obstacles.
+
+In order to combine the SVM classification with the edge detection, two strategies were proposed.
+The first strategy simply intersects the SVM classification results and edges to detect the nearest obstacle in the trajectory.
+This method show to be the safe option, by often falsely detecting an obstacle.
+Results show that the best strategy is to first detect all the edges and then classify these edges as being part of the road or not, using the SVM classifier.
+Only edges which are not considered part of the road remain.
+The maximum speed is then calculated based on the nearest edge in the trajectory of the car.
+
 \newpage

### 2016-06-15 Administratie/report/paper.tex
report update

@@
 For each frame in the dataset, a mask is given that shows the projected trajectory of the car in that frame.
-Intersections between this mask and the ''safe zone'' in the frame get calculated in order to find the maximum safe speed for each frame.
+They are used to derive the maximum speed to be able to brake safely without hitting an object.
+When an object is detected at a certain point on the mask, the intensity at that point gives the maximum safe driving speed.
 The edges of the edge detection are visualized as blue lines on figure \ref{visualization}.
@@
 
-Both strategies start by removing outliers from the road classification.
-The goal of this is to eliminate the effect that outliers would have when it's close to the car.
-A non-road block is considered an outlier when it has less than 4 non-road neighbors.
-This is repeated until all outliers are removed.
-
 For the first strategy, the edges of the area classified as road are computed.
@@
 For safety reasons there is an additional check.
-The above, left and right blocks of an edge should also be considered as non-road.
+The two blocks above, left and right of an edge should also be considered as non-road.
 When an edge lies on a part of the frame that is not classified as road, it is added to a collection.
 The intersections of the edges in the collection and the masks are then computed.
+Using the mask and the location of the intersection closest to the car, the maximum driving speed is derived.
 
@@
 The first bar of the histogram shows for every dataset the performance of road classification.
-Because of a crash in the first dataset there can only be traveled 317 meters (out of the total 1350 meters in this dataset).
 The second bar shows the distance traveled only relying on edge detection for each dataset.
@@
 	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
-\end{figure}
-
-As an extra, this strategy is also able to cope better with zebra crossings than the first strategy. This is shown in figure \ref{zebra}. In the first figure (\ref{zebra-1}) TODO
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
-  \caption{Zebra crossing filtered out by edge detection.}
-  \label{zebra-1}
-\end{subfigure}%
-\begin{subfigure}[t]{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
-  \caption{Zebra crossing classified as road.}
-  \label{zebra-2}
-\end{subfigure}
-\caption{Zebra crossing\label{zebra}}
 \end{figure}

### 2016-06-15 Administratie/report/paper.tex
Finishing up

--- 2016-06-15 Administratie/report/paper.tex
@@
 \usepackage[english]{babel}
+\usepackage{appendix}
 \usepackage{amssymb}
@@
 \begin{abstract}
-This is the abstract: The goal of this paper it to determine the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column.
+The goal of this paper it to determine the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column.
 We construct a system that combines SVM classification and edge detection to foresee possible obstacles on the given trajectory of the car.
@@
 
+All of these results for the classifier can be found in appendix \ref{appendix}.
+
 \subsubsection{Kernels}
@@
 \centering
-\includegraphics[width=\textwidth]{fig/RBF_Col_wi_RM.png}
+\includegraphics[width=\textwidth]{fig/RBF-COL-wi-RM.pdf}
 \caption{F-scores for color with road marks using RBF kernel}
@@
 
-\textbf{TODO} grafiek wo/wi RM
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{fig/16-wi-roadmarks.pdf}
+\caption{Comparision of F-scores for 16x16px block size with road marks using RBF kernel }
+\label{fig:16_wi_roadmarks}
+\end{figure}
 
@@
 
-TODO figure aanpassen
-\begin{figure}[h]
+\begin{figure}
 \centering
-\includegraphics[width=\textwidth]{fig/16_wi_roadmarks.png}
-\caption{Comparision of F-scores for 16x16px block size with road marks using RBF kernel }
-\label{fig:16_wi_roadmarks}
+\includegraphics[width=\textwidth]{fig/with-vs-without-roadmarks.pdf}
+\caption{F-scores for 16x16px blocks with color using RBF kernel}
+\label{fig:col-wi-rm}
 \end{figure}
@@
 
-\textbf{TODO} check if list not f***ed up.\\
 \noindent In retrospect, the optimal parameters for our classification problem are:
@@
 
+\newpage
+\appendix
+\section{F-scores of parameter iterations}
+\label{appendix}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth]{fig/RBF-LBP-wo-RM.pdf}
+\caption{F-scores for LBP without road marks using RBF kernel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth]{fig/RBF-COL-wo-RM.pdf}
+\caption{F-scores for color without road marks using RBF kernel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{fig/RBF-LBP-COL-wi-RM.pdf}
+\caption{F-scores for LBP and color without road marks using RBF kernel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{fig/RBF-LBP-wi-RM.pdf}
+\caption{F-scores for LBP with road marks using RBF kernel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{fig/RBF-COL-wi-RM.pdf}
+\caption{F-scores for color with road marks using RBF kernel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{fig/RBF-LBP-COL-wi-RM.pdf}
+\caption{F-scores for LBP and color with road marks using RBF kernel}
+\end{figure}
+
 \end{document}

### 2016-06-16 Administratie/report/paper.tex
Abstract tweak
Finale versie paper

@@
 The maximum speed is computed so that braking in time will avoid collision with those obstacles.
-TODO: hoe goed is dit systeem?
+Several strategies are proposed and compared and we conclude which strategy shows optimal performance for our application.
+
 \keywords{SVM classification, Local Binary Patterns, edge detection}
@@
 This speed detection would simulate the braking distance towards the nearest obstacle and inform the user.
+
 There are various ways to achieve this.
@@
 Basically, we want the hyperplane that maximizes the geometric distance to the closest data points (as shown in figure \ref{fig:hyperplane}).
+With this hyperplane, we can determine for any given data point on which side of the hyperplane it is located, i.e the class the data point belongs to.
 
@@
 In $v$-fold cross-validation, we first divide the training set into $v$ subsets of equal size.
-Sequentially one subset is tested using the classifier trained on the remaining $v − 1$ subsets.
+Sequentially one subset is tested using the classifier trained on the remaining $v-1$ subsets.
 Thus, each instance of the whole training set is predicted once so the cross-validation accuracy is the percentage of data which are correctly classified.
@@
 For a training set (a subset of the datasets available) each frame is divided into blocks.
-Each block must be manually labeled with the expected value for classification.
+Each block must be manually labelled with the expected value for classification.
 For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset ($i=0,10,...,40$).
-Figure \ref{fig:frame-training} shows a visualization of pre-labeled blocks in a frame.
+Figure \ref{fig:frame-training} shows a visualization of pre-labelled blocks in a frame.
 The classifier is trained using the label and calculated feature vector for each block in each frame of the training set.
@@
 \subsection{Road marks}
-The training dataset was labeled twice, once considering blocks containing road marks as part of the road class and once as part of the non-road class.
-Initially road marks were labeled as non-road, in order to get a good classification of road (i.e.
+The training dataset was labelled twice, once considering blocks containing road marks as part of the road class and once as part of the non-road class.
+Initially road marks were labelled as non-road, in order to get a good classification of road (i.e.
 different types of asphalt).
@@
 This makes up for a more difficult classification problem.
-To quantify this effect we apply classification both with and without road marks.
+To quantify this effect we tested and compared classification results both with and without road marks.
 
@@
 The trained classifier is then tested on the remaining datasets.
-
 All of these results for the classifier can be found in appendix \ref{appendix}.
@@
 As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}.
-Other combinations follow the same trend, although 32x32px sometimes performs slightly better.
-This is why we conclude the block size of 32x32px to be a more reliable choice.
+Other combinations follow the same trend.
+In practice, smaller blocks are more convenient for our application.
+Smaller blocks allow for a more precise estimation of road, and its edges.
+This is why we conclude the block size of 16x16px to be a more reliable choice.
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 The use of LBP represents no significant gains in order to get a viable classifier for road detection.
-If we compare results (block size 32x32px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
+If we compare results (block size 16x16px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
 Therefore, in the case of our classification problem, we do not add LBP to the feature vector.
+It should be mentioned that we only used the basic form of LBP, without any improvements such as Rotation Invariance \cite{Guo2010}.
 
@@
 \subsubsection{Road marks}
-Including road marks as part of the road gets better results for the classification problem.
+Including road marks as part of the road gets better results for the classification problem, as is shown in figure \ref{fig:wi-vs-wo-RM}.
 For our application this is more practical, because road marks should not be considered to be obstacles.
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \caption{F-scores for 16x16px blocks with color using RBF kernel}
-\label{fig:col-wi-rm}
+\label{fig:wi-vs-wo-RM}
 \end{figure}
@@
 This is because another type of road is used in this dataset, conflicting with the more common road type present in the other datasets.
-This distorts the classifier increases misclassifications.
+This distorts the classifier and increases misclassifications.
 
@@
 \item RBF kernel
-\item Block size 32x32px
+\item Block size 16x16px
 \item Using only color histogram values for the feature vectors
@@
 
-The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 32x32px in an image.
+The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 16x16px in an image.
 These results tend to get better when we include road marks as part of the road in the training data.
@@
 The first strategy simply intersects the SVM classification results and edges to detect the nearest obstacle in the trajectory.
-This method show to be the safe option, by often falsely detecting an obstacle.
+This method shows to be the safe option, by often falsely detecting an obstacle.
 Results show that the best strategy is to first detect all the edges and then classify these edges as being part of the road or not, using the SVM classifier.

### 2016-06-16 maxSpeed.cpp
Removed 30-frame limit

@@ -351,15 +351,8 @@ int main(int argc, char** argv){
         iss >> speed;
         speeds.push_back(speed);
     }
-    vector<Mat> masks_all = read_images(dirDataset,"mask%05d.png");
-    vector<Mat> frames_all = read_images(dirDataset,"frame%05d.png");
-
-    vector<Mat> masks;
-    vector<Mat> frames;
-    for(int i=0; i < 30; i++) {
-        masks.push_back(masks_all[i]);
-        frames.push_back(frames_all[i]);
-    }
+    vector<Mat> masks = read_images(dirDataset,"mask%05d.png");
+    vector<Mat> frames = read_images(dirDataset,"frame%05d.png");
 
 
     vector<Mat> roads = detectLines(masks,frames);

### 2016-06-16 maxSpeed.cpp
maxSpeed: outlier removal disabled

@@ -102,7 +102,8 @@ void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames
 
 
         //rode outliers uithalen
-        bool changedOutlier = true;
+        // --> Disabled
+        /*bool changedOutlier = true;
         while(changedOutlier) {
             changedOutlier = false;
             for (int z = blksInWidth; z < roadRegions[i].size() - blksInWidth; z++) {
@@ -135,7 +136,7 @@ void showMaxSpeed(vector<Mat> & masks, vector<Mat> & roads, vector<Mat> & frames
                     }
                 }
             }
-        }
+        }*/
 
         // Add non-road rectangles to the canny output
         // --> Disabled

### 2016-06-17 Dataset/evaluate.py
Code gekuist: verwijderen van overbodig commentaar, code en output. Map indienen met geteste binaries

@@ -26,6 +26,3 @@ print("You travelled %f metres in %f seconds."%(distance,time))
 #plt.plot(speeds)
 #plt.ylim([0,30])
 #plt.show()
-	
-
-

### 2016-06-17 Headers/io.hpp

@@ -15,10 +15,10 @@ using namespace cv;
 
 namespace io {
 
-	/** Constants **/
+	/** Directories and data filenames **/
 	const string dirOutput = "output/";
-	const string dirOutputLogging = dirOutput+"Logging/";
-	const string dirTrainingsdata = "Trainingsdata/";
+	const string dirOutputLogging = "Logging/";
+	const string dirData = "Data/";
 	const string labelsPostfix = "_labels.csv";
 	const string lbpStr = "_lbp";
 	const string colorStr = "_color";
@@ -27,18 +27,19 @@ namespace io {
 	const int KEY_ESCAPE = 537919515;
 	
 	/** Maxspeed training and run parameters**/
+    const int blkSize = 16; // 8,16,32
+    const bool useLBP = false; // Make sure to use either color, LBP or both
+    const bool useColor = true;
+    const bool includeMarks = true;
+    const bool trainAuto = false;
+
 	const string datasetFolders[4] = { "Dataset/01/", "Dataset/02/","Dataset/03/", "Dataset/04/" };
     const int datasets[] = {1,2,3,4}; //{1,2,3,4};
     const int frameInterval = 10;
     const int frameStopIdx = 50;
     const int innerMargin = 64;
-    const int blkSize = 16; // 8,16,32
     const int lbpRadius = 1;
-    const bool includeMarks = true;
-    const bool trainAuto = true;
-    const bool useLBP = false; // Make sure to use either color, LBP or both
-    const bool useColor = true;
-
+    
     /** SVM parameters for when not training automatically **/
     const double C = 0.1;
     const double gamma = 0.00001;
@@ -162,61 +163,14 @@ namespace io {
 	    frameName = string(buffer) + (includeMarks?marksStr:"");
 	}
 
-	void readTrainingsdata(vector<short>& datasets, Mat& labelsMat, Mat& trainingsMat){
+	void readTrainingsdata(string frameName, Mat& labelsMat, Mat& trainingsMat, bool useLBP, bool useColor){
 		vector<short> labels;
 		vector<vector<int> > featureVectors;
 
-		for(int s=0;s<datasets.size();s++) {
-			for(int f=0; f<=200; f+=5) {
-			    char buffer[36];
[…]

### 2016-06-17 Headers/lbpfeaturevector.hpp

@@ -177,7 +177,7 @@ public:
 	*/
 	void processFrame(string fnFrame, Mat& img, Mat& featVectors, bool useColor=io::useColor, bool useLBP=io::useLBP){
 
-    	cout<<"Start processFrame "<<fnFrame+(useLBP?io::lbpStr:"")+(useColor?io::colorStr:"")<<endl;
+    	// cout<<"Start processFrame "<<fnFrame+(useLBP?io::lbpStr:"")+(useColor?io::colorStr:"")<<endl;
     	if(!useColor && !useLBP){
     		cerr<<"You are using neither color or LBP for featurvectors, aka nothing!"<<endl;
     		throw;

### 2016-06-17 Headers/svm.hpp

@@ -35,7 +35,7 @@ public:
 	    // svm->setKernel(SVM::LINEAR);
 	    svm->setTermCriteria(TermCriteria(TermCriteria::MAX_ITER, 100, 1e-6));
 	    Ptr<TrainData> trainData_ptr = TrainData::create(trainingsMat, ROW_SAMPLE , labelsMat);
-		cout << "training the SVM... "<<(trainAuto? "(automatically, this could take a while...)" : "") << endl;
+		cout << "Training the SVM... "<<(trainAuto? "(automatically, this could take a while...)" : "") << endl;
 	    if(trainAuto){
 		    /* Automatic training */
 		    ParamGrid emptyGrid(0,0,0);
@@ -150,7 +150,8 @@ void my_svm::test(int min = 1, int max = 4, int skip = 5, int number = 5) {
     for(int ds=min; ds<=max; ds++) {
         char dsnumber[36];
         sprintf(dsnumber, "%02d", ds);
-        for(int f=5; f<=(5+number*skip); f+=skip) {
+        // for(int f=5; f<=(5+number*skip); f+=skip) {
+        for(int f=0; f<=(number*skip); f+=skip) {
             char number[36];
             sprintf(number, "%05d", f);
             // Show decision regions by the SVM
@@ -189,7 +190,7 @@ void my_svm::test(int min = 1, int max = 4, int skip = 5, int number = 5) {
 }
 
 void my_svm::printParams(ostream& os){
-	os<<"--- Params ---"<<endl;
+	os<<"--- Parameters ---"<<endl;
 	os<<"C:\t\t"<<				svm->getC()<<endl;
 	os<<"Class weights:\t"<<	svm->getClassWeights()<<endl;
 	os<<"Coef0:\t\t"<<			svm->getCoef0()<<endl;

### 2016-06-17 run_maxSpeed.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+rm -rf CMake*
+openCvCmake.bash maxSpeed
+./maxSpeed $1
+subl ./Dataset/evaluate.py
+nemo ./outputframes/$1

### 2016-06-17 run_svm.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+rm -rf CMake*
+openCvCmake.bash svm
+./svm

### 2016-06-17 trainingsdataFTT.cpp

@@ -1,346 +1,0 @@
-#include <string>
-#include <stdio.h>
-// #include <time.h>
-#include <set>
-#include "Headers/io.hpp"
-#include "Headers/lbpfeaturevector.hpp"
-#include "Headers/svm.hpp"
-
-using namespace std;
-using namespace cv;
-
-void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
-	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
-	if(io::file_exists(fnLabels)){
-		io::readFrameLabels(fnLabels,isRoad);
-	}else{
-		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
-		isRoad = vector<bool>(totalBlocks);
-
-		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
-			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
-	}
-}
-
-static void onMouse( int event, int x, int y, int, void* param){
-    if( event != EVENT_LBUTTONDOWN )
-        return;
-
-    int* coordinates = (int*) param;
-    coordinates[0] = x;
-    coordinates[1] = y;
-}
-
-void fileToFrameName(string fnFrame, string& frameName){
-	size_t pos1 = fnFrame.find_last_of("/")+1;
-	if(pos1 == string::npos) pos1=0;
-	size_t pos2 = fnFrame.find_last_of(".");
-	if(pos2 == string::npos) pos2=fnFrame.size();
-	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
-}
-
-void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
-	/* Initial configuration */
-    ostringstream strBldr; // Stringbuilder
-	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
-	imWidth = img.cols;
-	imHeight = img.rows;
-	outerMargin = fv.getOuterMargin();
-	innerMargin = fv.getInnerMargin();
-	blkX = blkY = outerMargin;
-	blkSize = fv.getBlkSize();
-
-	/* Draw initial block labeling */
-	for(int blkIdx=0; 
[…]

### 2016-06-17 trainingsdataTFT.cpp

@@ -1,346 +1,0 @@
-#include <string>
-#include <stdio.h>
-// #include <time.h>
-#include <set>
-#include "Headers/io.hpp"
-#include "Headers/lbpfeaturevector.hpp"
-#include "Headers/svm.hpp"
-
-using namespace std;
-using namespace cv;
-
-void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
-	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
-	if(io::file_exists(fnLabels)){
-		io::readFrameLabels(fnLabels,isRoad);
-	}else{
-		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
-		isRoad = vector<bool>(totalBlocks);
-
-		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
-			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
-	}
-}
-
-static void onMouse( int event, int x, int y, int, void* param){
-    if( event != EVENT_LBUTTONDOWN )
-        return;
-
-    int* coordinates = (int*) param;
-    coordinates[0] = x;
-    coordinates[1] = y;
-}
-
-void fileToFrameName(string fnFrame, string& frameName){
-	size_t pos1 = fnFrame.find_last_of("/")+1;
-	if(pos1 == string::npos) pos1=0;
-	size_t pos2 = fnFrame.find_last_of(".");
-	if(pos2 == string::npos) pos2=fnFrame.size();
-	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
-}
-
-void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
-	/* Initial configuration */
-    ostringstream strBldr; // Stringbuilder
-	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
-	imWidth = img.cols;
-	imHeight = img.rows;
-	outerMargin = fv.getOuterMargin();
-	innerMargin = fv.getInnerMargin();
-	blkX = blkY = outerMargin;
-	blkSize = fv.getBlkSize();
-
-	/* Draw initial block labeling */
-	for(int blkIdx=0; 
[…]

### 2016-06-17 trainingsdataTTT.cpp

@@ -1,346 +1,0 @@
-#include <string>
-#include <stdio.h>
-// #include <time.h>
-#include <set>
-#include "Headers/io.hpp"
-#include "Headers/lbpfeaturevector.hpp"
-#include "Headers/svm.hpp"
-
-using namespace std;
-using namespace cv;
-
-void generateLabels(string frameName, vector<bool>& isRoad, int totalBlocks){
-	string fnLabels(io::dirOutput+frameName+io::labelsPostfix);
-	if(io::file_exists(fnLabels)){
-		io::readFrameLabels(fnLabels,isRoad);
-	}else{
-		int isRoadThreshold = (totalBlocks-1)*0.6; // First/upper 60% of blocks is likely not road
-		isRoad = vector<bool>(totalBlocks);
-
-		for(int blkIdx=0;blkIdx<totalBlocks;blkIdx++)
-			isRoad[blkIdx] = (blkIdx>isRoadThreshold);
-	}
-}
-
-static void onMouse( int event, int x, int y, int, void* param){
-    if( event != EVENT_LBUTTONDOWN )
-        return;
-
-    int* coordinates = (int*) param;
-    coordinates[0] = x;
-    coordinates[1] = y;
-}
-
-void fileToFrameName(string fnFrame, string& frameName){
-	size_t pos1 = fnFrame.find_last_of("/")+1;
-	if(pos1 == string::npos) pos1=0;
-	size_t pos2 = fnFrame.find_last_of(".");
-	if(pos2 == string::npos) pos2=fnFrame.size();
-	frameName = fnFrame.substr(pos1,pos2-pos1); // Remove folder prefixes and file extension
-}
-
-void guiLabeling(LbpFeatureVector& fv, Mat& img, vector<bool>& isRoad){
-	/* Initial configuration */
-    ostringstream strBldr; // Stringbuilder
-	int blkX,blkY,imWidth,imHeight,outerMargin,innerMargin,blkSize; // Necessary parameters for block coordinates and size
-	imWidth = img.cols;
-	imHeight = img.rows;
-	outerMargin = fv.getOuterMargin();
-	innerMargin = fv.getInnerMargin();
-	blkX = blkY = outerMargin;
-	blkSize = fv.getBlkSize();
-
-	/* Draw initial block labeling */
-	for(int blkIdx=0; 
[…]

### 2016-06-17 INDIENEN/Headers/LineDetection.hpp
ReadMe

@@ -1,0 +1,98 @@
+#include <cstdio>
+#include <string>
+#include <vector>
+#include "opencv2/opencv.hpp"
+#include "opencv2/highgui.hpp"
+#include "opencv2/core/core.hpp"
+#include "opencv2/imgproc/imgproc.hpp"
+
+#define PI 3.1415926
+
+
+using namespace std;
+using namespace cv;
+
+
+class LineDetection {
+
+public:
+    Mat getLinesFromImage(Mat &image) {
+        return getLinesFromImage(image, false);
+    }
+
+    Mat getLinesFromImage(Mat &image, bool showSteps) {
+
+        // parameters
+        int cannyLowerThreshold = 45; // 40
+        int cannyUpperThreshold = cannyLowerThreshold * 2;
+        int houghPMinLineLength = 7;
+        int houghPGap = 10;
+        int houghPMinVote = 7;
+        double houghPDeltaRho = 1;
+        double houghPDeltaTheta = PI / 180;
+
+        // set the ROI for the image
+        Rect roi(0, image.cols / 4, image.cols - 1, image.rows - image.cols / 4);
+        Mat imgROI = image(roi);
+
+
+        Mat tmp;
+        Size size;
+        size.width = 100; // 100
+        size.height = 30; // 20
+        Mat kernel = getStructuringElement(MORPH_ELLIPSE, size);
+        erode(imgROI, tmp, kernel);
+
+        size.width = 85; // 85
+        size.height = 25; // 18
+        kernel = getStructuringElement(MORPH_ELLIPSE, size);
+        dilate(tmp, tmp, kernel);
+
+
+        // Canny algorithm
+        Mat contours;
+        Mat result(imgROI.size(), CV_8U, Scalar(0));
+        Canny(tmp, contours, cannyLowerThreshold, cannyUpperThreshold, 3);
+        Mat contoursInv;
+        contours.copyTo(result);
+
+        // Do probabilistic Hough
+        vector<Vec4i> lines;
+
+        HoughLinesP(contours,lines,houghPDeltaRho,houghPDeltaTheta,houghPMinVote, houghPMinLineLength, houghPGap);
+        //drawDetectedLines(result, lines);
+
+
+        if (showSteps) {
+            namedWindow("Input", CV_WINDOW_AUTOSIZE);
+            imshow("Input", tmp);
+            namedWindow("Canny", CV_WINDOW_AUTOSIZE);
+            imshow("Canny", contours);
[…]

### 2016-06-17 INDIENEN/Headers/io.hpp

@@ -1,0 +1,237 @@
+#ifndef IO_
+#define IO_
+
+#include <string>
+#include <fstream>
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <unistd.h>
+#include <ctime>
+#include "opencv2/opencv.hpp"
+#include "opencv2/highgui.hpp"
+
+using namespace std;
+using namespace cv;
+
+namespace io {
+
+	/** Directories and data filenames **/
+	const string dirOutput = "output/";
+	const string dirOutputLogging = "Logging/";
+	const string dirData = "Data/";
+	const string labelsPostfix = "_labels.csv";
+	const string lbpStr = "_lbp";
+	const string colorStr = "_color";
+	const string marksStr = "_marks";
+	const string featVecsPostfix = "_featurevectors.csv";
+	const int KEY_ESCAPE = 537919515;
+	
+	/** Maxspeed training and run parameters**/
+    const int blkSize = 16; // 8,16,32
+    const bool useLBP = false; // Make sure to use either color, LBP or both
+    const bool useColor = true;
+    const bool includeMarks = true;
+    const bool trainAuto = false;
+
+	const string datasetFolders[4] = { "Dataset/01/", "Dataset/02/","Dataset/03/", "Dataset/04/" };
+    const int datasets[] = {1,2,3,4}; //{1,2,3,4};
+    const int frameInterval = 10;
+    const int frameStopIdx = 50;
+    const int innerMargin = 64;
+    const int lbpRadius = 1;
+    
+    /** SVM parameters for when not training automatically **/
+    const double C = 0.1;
+    const double gamma = 0.00001;
+
+	/** Shown image properties and helper values **/
+	int imX=50,imY=50
+		,shownImages=0
+		,consecutiveImages=6
+		,normalSize=325;
+
+	map<string,vector<int> > imgPos;
+
+	void showImage(string windowName,Mat& img, bool resize=false){
+	    int windowFlag, xSize, ySize, margin=3;
+	    if(resize){
+	        windowFlag = WINDOW_NORMAL;
+	        xSize = ySize = normalSize;
+	    } else {
+	        windowFlag = WINDOW_AUTOSIZE;
+	        xSize = img.cols;
[…]

### 2016-06-17 INDIENEN/Headers/lbpfeaturevector.hpp

@@ -39,6 +39,10 @@ public:
 	*/
 	template <typename _Tp>
 	void _LBP(const Mat& src, Mat& dst, int r, int x, int y, int winSize) {
+	    // if( x<r || y<r || x+winSize+r>src.cols || y+winSize+r>src.rows){
+	    //     cerr << "Error in _LBP : source coordinates and window size make radius exceed image size" << endl;
+	    //     throw;
+	    // }
 	    dst = Mat::zeros(winSize, winSize, CV_8UC(src.channels()));
 	    vector<Mat> srcChnls(src.channels()) , dstChnls(src.channels());
 	    split(src,srcChnls);

### 2016-06-17 INDIENEN/Headers/svm.hpp

@@ -1,0 +1,202 @@
+#ifndef SVM_
+#define SVM_
+
+#include <string>
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <unistd.h>
+#include "opencv2/opencv.hpp"
+#include "opencv2/highgui.hpp"
+#include <opencv2/ml.hpp>
+#include "io.hpp"
+#include "lbpfeaturevector.hpp"
+
+using namespace std;
+using namespace cv;
+using namespace cv::ml;
+
+class my_svm {	
+private:
+	Ptr<SVM> svm;
+	// Mat labelsMat;
+	// Mat trainingsMat;
+	Mat confusion;
+	double precision;
+	double recall;
+	double accuracy;
+	double true_negative;
+	double F;
+
+public:
+	my_svm(Mat& labelsMat, Mat& trainingsMat, bool trainAuto=false) {
+		svm = SVM::create();
+	    svm->setType(SVM::C_SVC);
+	    svm->setKernel(SVM::RBF);
+	    // svm->setKernel(SVM::LINEAR);
+	    svm->setTermCriteria(TermCriteria(TermCriteria::MAX_ITER, 100, 1e-6));
+	    Ptr<TrainData> trainData_ptr = TrainData::create(trainingsMat, ROW_SAMPLE , labelsMat);
+		cout << "Training the SVM... "<<(trainAuto? "(automatically, this could take a while...)" : "") << endl;
+	    if(trainAuto){
+		    /* Automatic training */
+		    ParamGrid emptyGrid(0,0,0);
+		    svm->trainAuto(trainData_ptr,10,SVM::getDefaultGrid(SVM::C),SVM::getDefaultGrid(SVM::GAMMA),emptyGrid,emptyGrid,emptyGrid,emptyGrid,true);
+		}else{
+		    /* Hardcoded params */
+		    svm->setC(io::C);
+		    svm->setCoef0(0);
+		    svm->setDegree(0);
+		    svm->setGamma(io::gamma);
+		    svm->setNu(0);
+		    svm->setP(0);
+		    svm->train(trainData_ptr);
+		}
+		cout << "SVM trained" << endl;
+		confusion = Mat::zeros(2,2, CV_32S);
+		precision = 0.0;
+		recall = 0.0;
+		accuracy = 0.0;
+		true_negative = 0.0;
+		F = 0.0;
+
+	    printParams(cout);
+	}
+
+	Ptr<SVM> get_svm();
+	void calculateScores(Mat&, Mat&);
[…]

### 2016-06-17 INDIENEN/maxSpeed.cpp

@@ -306,6 +306,13 @@ int main(int argc, char** argv){
     }
     string dirDataset(argv[1]);
     string results;
+
+    // Remove ending (back)slash
+    const char* delim = "/\\";
+    size_t pos = dirDataset.find_last_of(delim);
+    if(pos==(dirDataset.length()-1)){
+        dirDataset = dirDataset.substr(0,pos);
+    }
     
     vector<double> speeds;
     string line;
@@ -334,12 +341,8 @@ int main(int argc, char** argv){
     cout << "Road detection done." << endl;
     
     string subdirOutFra = dirDataset;
-    const char* delim = "/\\";
-    size_t pos = subdirOutFra.find_last_of(delim);
+    pos = subdirOutFra.find_last_of(delim);
     if(pos != string::npos){
-        if(pos==subdirOutFra.length()-1){
-            pos = subdirOutFra.find_last_of(delim,pos);
-        }
         subdirOutFra = subdirOutFra.substr(pos+1);
     }
     string dirOutputFrames = "outputframes/"+subdirOutFra;
